---
title: "02b - Case Data Preparation"
author: "Prince Oppong Boakye"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: show
    theme: flatly
---

# Case Data Preparation

Prepares COVID-19 case rates for all study periods.
- **NYC**: weekly MODZCTA case rates → daily interpolation → borough spillover
- **Utah**: NYT daily cumulative county counts → daily new cases → rate per 100k → neighbor spillover

Output: `tables/case_rates.csv` per month with columns:
`geo_id`, `date`, `sim_day`, `case_rate`, `case_rate_spillover`, `region`.

```{r setup, include=FALSE}
source(here::here("helpers/_utils.R"))
load_pipeline()
library(data.table)
library(lubridate)
```

```{r gamma_t_helper}
# Testing-adjustment γ_t (Meyer & Held 2017 §4.3; Slater et al. 2025 §3.4).
# γ_t = (positivity_t / median(positivity))^(-0.5), clamped to ±log(3).
# If data/us_CovidCases/testing/<state>_positivity.csv (date, positivity) is
# absent, returns 0 (offset unchanged); the column is always present for 08a.
load_log_gamma_t <- function(state, dates) {
  fp <- here::here("data", "us_CovidCases", "testing",
                   sprintf("%s_positivity.csv", state))
  out <- data.table(date = as.Date(dates), log_gamma_t = 0)
  if (!file.exists(fp)) return(out)
  pos <- fread(fp)
  pos[, date := as.Date(date)]
  pos[, positivity := pmax(as.numeric(positivity), 1e-4)]
  setorder(pos, date)
  pos[, positivity := zoo::rollmean(positivity, k = 7, fill = NA, align = "center")]
  pos[is.na(positivity), positivity := median(pos$positivity, na.rm = TRUE)]
  med <- median(pos$positivity, na.rm = TRUE)
  pos[, log_gamma_t := pmin(pmax(-0.5 * (log(positivity) - log(med)),
                                 -log(3)), log(3))]
  out <- merge(out[, .(date)], pos[, .(date, log_gamma_t)],
               by = "date", all.x = TRUE)
  out[is.na(log_gamma_t), log_gamma_t := 0]
  out
}
```

# Load NYC MODZCTA Data

```{r load_data}
NY_WAVES <- WAVE_RUNS[state == "NY"]
UT_WAVES <- WAVE_RUNS[state == "UT"]
```

```{r nyc_load, eval = nrow(NY_WAVES) > 0}
## --- Weekly case rates (wide: week_ending x CASERATE_<modzcta>) ---
nyc_raw <- fread(CONFIG$nyc_caserate_path)
modzcta_cols <- grep("^CASERATE_[0-9]+$", names(nyc_raw), value = TRUE)

nyc_long <- melt(nyc_raw, id.vars = "week_ending", measure.vars = modzcta_cols,
                 variable.name = "modzcta_col", value.name = "case_rate")
nyc_long[, modzcta := gsub("CASERATE_", "", modzcta_col)]
nyc_long[, modzcta_col := NULL]
nyc_long[, week_ending_date := mdy(week_ending)]
nyc_long <- nyc_long[!is.na(week_ending_date) & !is.na(case_rate)]


## --- CBG-to-MODZCTA crosswalk (for borough assignment) ---
cbg_modzcta <- fread(CONFIG$nyc_crosswalk_path)
cbg_modzcta[, modzcta := as.character(modzcta)]
saveRDS(cbg_modzcta, file.path(CONFIG$cache_dir, "cbg_modzcta_crosswalk_NY.rds"))

# Unique MODZCTA -> borough mapping (take first if multiple)
modzcta_borough <- cbg_modzcta[, .(borough = borough[1]), by = modzcta]
```

# Process Each Period

```{r process_periods, eval = nrow(NY_WAVES) > 0}
process_nyc_wave <- function(wave_id, wave_months) {
  wdr <- get_wave_date_range(wave_months)
  out_dir <- get_wave_dir("NY", wave_id)
  dir.create(file.path(out_dir, "tables"), recursive = TRUE, showWarnings = FALSE)

  ## --- Filter weeks overlapping this wave (with 6-day pad on each side) ---
  relevant <- nyc_long[week_ending_date >= (wdr$start - 6) &
                       week_ending_date <= (wdr$end + 6)]

  ## --- Weekly midpoints for smooth interpolation across the full 90-day window ---
  midpoints <- relevant[, .(mid_date = week_ending_date - 3L, modzcta, case_rate)]
  midpoints <- midpoints[, .(case_rate = mean(case_rate, na.rm = TRUE)),
                         by = .(modzcta, mid_date)]
  setorder(midpoints, modzcta, mid_date)

  all_dates <- seq(wdr$start, wdr$end, by = "day")
  daily <- rbindlist(lapply(unique(midpoints$modzcta), function(mz) {
    mz_data <- midpoints[modzcta == mz]
    if (nrow(mz_data) < 2L) {
      return(data.table(modzcta = mz, date = all_dates,
                        case_rate = mz_data$case_rate[1]))
    }
    interp <- approx(x = as.numeric(mz_data$mid_date),
                     y = mz_data$case_rate,
                     xout = as.numeric(all_dates),
                     rule = 2)  # extrapolate nearest at edges
    data.table(modzcta = mz, date = all_dates, case_rate = pmax(interp$y, 0))
  }))
  daily[, sim_day := as.integer(date - wdr$start) + 1L]
  # Tag each day with its source month for diagnostic faceting downstream.
  daily[, source_month := format(date, "%Y-%m")]

  ## --- Borough assignment + spillover ---
  daily[, borough := modzcta_borough[.SD, borough, on = .(modzcta)]]
  daily[is.na(borough), borough := "Unknown"]

  borough_rates <- daily[, .(borough_mean = mean(case_rate, na.rm = TRUE)),
                         by = .(date, borough)]
  daily[borough_rates, borough_mean := i.borough_mean, on = .(date, borough)]

  alpha <- CONFIG$spillover_alpha
  daily[, case_rate_spillover := alpha * case_rate + (1 - alpha) * borough_mean]
  daily[is.na(case_rate_spillover), case_rate_spillover := case_rate]

  ## --- Testing-adjustment γ_t ---
  gamma_dt <- load_log_gamma_t("NY", unique(daily$date))
  daily <- merge(daily, gamma_dt, by = "date", all.x = TRUE)
  daily[is.na(log_gamma_t), log_gamma_t := 0]

  ## --- Output (90 continuous sim_days) ---
  out <- daily[, .(geo_id = modzcta, date, sim_day, source_month,
                   case_rate, case_rate_spillover, borough, log_gamma_t)]
  fwrite(out, file.path(out_dir, "tables", "case_rates.csv"))
  out
}

results <- lapply(seq_len(nrow(NY_WAVES)), function(i) {
  process_nyc_wave(NY_WAVES$wave_id[i], NY_WAVES$wave_months[[i]])
})
names(results) <- NY_WAVES$wave_id
```

# Utah Case Data (NYT Daily County Counts)

```{r ut_load, eval = nrow(UT_WAVES) > 0}
# Load NYT county-level daily cumulative counts for all years touched by waves
ut_years <- unique(unlist(lapply(UT_WAVES$wave_months, function(ms) substr(ms, 1, 4))))
nyt_raw <- rbindlist(lapply(ut_years, function(y) {
  fp <- CONFIG$nyt_counties_files[[y]]
  if (!file.exists(fp)) { return(NULL) }
  fread(fp)
}))
# Filter to Utah
ut_raw <- nyt_raw[state == "Utah" & !is.na(fips)]
ut_raw[, fips := sprintf("%05d", as.integer(fips))]
ut_raw[, date := as.Date(date)]
setorder(ut_raw, fips, date)
# Compute daily new cases from cumulative
ut_raw[, new_cases := cases - data.table::shift(cases, 1, fill = 0), by = fips]
ut_raw[new_cases < 0, new_cases := 0]  # handle data corrections

# Redistribute batch-reported cases (Nash et al. 2023; Simpson et al. 2022)
ut_raw <- redistribute_batch_cases(ut_raw[date >= as.Date("2020-10-01")],
                                   id_col = "fips", cases_col = "new_cases") |>
  rbind(ut_raw[date < as.Date("2020-10-01")]) |>
  setorder(fips, date)


# County population for rate calculation (from grid population CSV, year 2020 as baseline)
# We'll use the CDC surveillance data to cross-check counts
ut_cdc <- fread(CONFIG$ut_county_cases_path)
ut_cdc <- ut_cdc[!is.na(county_fips_code)]
```

```{r ut_process, eval = nrow(UT_WAVES) > 0}
process_ut_wave <- function(wave_id, wave_months) {
  wdr <- get_wave_date_range(wave_months)
  out_dir <- get_wave_dir("UT", wave_id)
  dir.create(file.path(out_dir, "tables"), recursive = TRUE, showWarnings = FALSE)

  # Filter NYT daily data to the full 90-day wave window
  daily <- ut_raw[date >= wdr$start & date <= wdr$end]

  # County population from wave's ACS year
  pop_year <- UT_WAVES$acs_year[UT_WAVES$wave_id == wave_id][1]
  pop_file <- file.path(CONFIG$grids_dir, sprintf("population_UT_%s.csv", pop_year))
  grid_pop <- fread(pop_file)
  grid_sf_path <- file.path(CONFIG$grids_dir, "grid_100m_UT.gpkg")

  gsf <- sf::st_read(grid_sf_path, quiet = TRUE)
  gdt <- data.table(grid_id = gsf$grid_id, fips_5 = gsf$fips_5)
  gdt <- merge(gdt, grid_pop[, .(grid_id, population)], by = "grid_id", all.x = TRUE)
  county_pop <- gdt[, .(population = sum(population, na.rm = TRUE)), by = fips_5]

  # Merge population and compute rate per 100k
  daily <- merge(daily, county_pop, by.x = "fips", by.y = "fips_5", all.x = TRUE)
  daily[is.na(population) | population < 100, population := 100]
  daily[, case_rate := new_cases / population * 1e5]
  daily[, sim_day := as.integer(date - wdr$start) + 1L]
  daily[, source_month := format(date, "%Y-%m")]

  # Region assignment
  daily[, region := county]

  # Statewide spillover
  state_mean <- daily[, .(state_mean = mean(case_rate, na.rm = TRUE)), by = date]
  daily <- merge(daily, state_mean, by = "date", all.x = TRUE)
  alpha <- CONFIG$spillover_alpha
  daily[, case_rate_spillover := alpha * case_rate + (1 - alpha) * state_mean]
  daily[is.na(case_rate_spillover), case_rate_spillover := case_rate]

  ## --- Testing-adjustment γ_t ---
  gamma_dt <- load_log_gamma_t("UT", unique(daily$date))
  daily <- merge(daily, gamma_dt, by = "date", all.x = TRUE)
  daily[is.na(log_gamma_t), log_gamma_t := 0]

  out <- daily[, .(geo_id = fips, date, sim_day, source_month,
                   new_cases, case_rate, case_rate_spillover, region, log_gamma_t)]
  fwrite(out, file.path(out_dir, "tables", "case_rates.csv"))
  out
}

ut_results <- lapply(seq_len(nrow(UT_WAVES)), function(i) {
  process_ut_wave(UT_WAVES$wave_id[i], UT_WAVES$wave_months[[i]])
})
names(ut_results) <- UT_WAVES$wave_id
```

# Summary

```{r summary}
all_results <- c(
  if (exists("results")) results else list(),
  if (exists("ut_results")) ut_results else list()
)
for (wid in names(all_results)) {
  dt <- all_results[[wid]]
}
```
