---
title: "03 - Advan Weekly Patterns Plus: Batch Extraction"
author: "Prince Oppong Boakye"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: show
    theme: flatly
---

# Advan Weekly Patterns Plus: Batch Extraction

Reads weekly parquet files (`data/advan_weekly_plus/`), aggregates each
weekly POI row into a per-(POI, wave_month) row, and writes the same
downstream artefacts the old monthly pipeline produced:

- `grid_visits.feather` (per-grid wave-aggregated POI metrics)
- `daily_stats.csv`     (per-day visits and dwell, 90 days)
- `od_home_cbg.feather` (POI ↔ home-CBG visitor pairs)
- `trip_chains.feather` (brand co-visitation, day & week scope)
- `daytime_visitors.feather` (POI-level daytime presence)

Schema migration notes (Weekly Plus vs old Monthly Patterns):

- `RAW_VISIT_COUNTS`           -> `VISIT_COUNTS`
- `RAW_VISITOR_COUNTS`         -> `VISITOR_COUNTS`
- `BRANDS`                     -> `BRAND`
- `RELATED_SAME_MONTH_BRAND`   -> `RELATED_SAME_WEEK_BRAND`
- `POPULARITY_BY_DAY`          -> derived from `VISITS_BY_DAY` (7-vec/wk)
- `POPULARITY_BY_HOUR`         -> derived from `VISITS_BY_EACH_HOUR`
                                  (168-vec/wk, summed 7×24 -> 24-vec)
- `DEVICE_TYPE`: was a label; now JSON `{"ios":N,"android":N}` -> split
  visits proportionally instead of exclusively by device label.
- `NORMALIZED_VISITS_BY_STATE_SCALING`: not present.  We use raw
  `VISIT_COUNTS` for both `visits_total` and `visits_norm`. (Panel
  correction is now downstream, in 04 / 08a, via observed device share.)
- `ENCLOSED`                   -> not present; replaced by NAICS-based
  `enclosed_mod` proxy (1.0 indoor / 0.5 mixed / 0.0 outdoor) via the
  `naics_lookup` table. POIs without a sub-category match default to
  1.0 (indoor).
- `WKT_AREA_SQ_METERS`         -> not present; replaced by spatial join
  to Microsoft Building Footprints in `build_poi_area_cache()` (helper),
  yielding `poi_area_sqm`. `visit_density` and `contact_density_hourly`
  are then derived here. Match rate is logged; <80 % triggers a warning.
- `OPEN_HOURS`                 -> not present; replaced by NAICS-typical
  `effective_hours` proxy (8-24 h/day) in `naics_lookup`. Default 12 h
  if no sub-category match.
- `OPEN_DATE` / `CLOSE_DATE`   -> new fields; weeks outside the POI's
  active window are dropped in `process_wave`.
- `IS_DISTRIBUTOR`             -> new field; rows with `is_distributor=1`
  are dropped in `process_wave`.

Each weekly partition covers one Mon-Sun window (`DATE_RANGE_START`
is a Monday). For wave-level aggregation we sum scalar visits across
all weeks falling in the wave date range, and merge JSON dicts (visitor
home CBGs, brand co-visits, dwell buckets, device type, daytime CBGs)
by summing values for matching keys.

```{r setup, include=FALSE}
source(here::here("helpers/_utils.R"))
load_pipeline()
library(data.table)
library(DBI)
library(duckdb)
library(sf)
library(arrow)
library(lubridate)
library(jsonlite)
```

# Category Risk Table

```{r category_risk}
# Single SUB_CATEGORY (6-digit NAICS) lookup with three columns:
#   category_risk    — relative transmission-risk multiplier
#   enclosed_mod     — 1.0 indoor, 0.5 mixed/semi-outdoor, 0.0 outdoor
#                      (proxy for the missing ENCLOSED flag in Weekly+)
#   effective_hours  — typical operating hours per day
#                      (proxy for the missing OPEN_HOURS field)
# A row uses NA for any field that is unknown for that sub-category.
naics_lookup <- data.table(
  sub_category = c(
    "Religious Organizations", "Drinking Places (Alcoholic Beverages)",
    "Full-Service Restaurants", "Cafeterias, Grill Buffets, and Buffets",
    "Fitness and Recreational Sports Centers", "Personal Care Services",
    "Barber Shops", "Beauty Salons", "Nail Salons",
    "Offices of Dentists", "Offices of Physicians",
    "Offices of Other Health Practitioners",
    "Child Day Care Services", "Elementary and Secondary Schools",
    "Colleges, Universities, and Professional Schools",
    "Museums, Historical Sites, and Similar Institutions",
    "Bowling Centers", "All Other Amusement and Recreation Industries",
    "Limited-Service Restaurants", "Snack and Nonalcoholic Beverage Bars",
    "Specialty Food Stores", "Meat Markets",
    "Other Schools and Instruction", "Clothing Stores",
    "Shoe Stores", "Jewelry Stores", "Sporting Goods Stores",
    "Drycleaning and Laundry Services",
    "Supermarkets and Other Grocery (except Convenience) Stores",
    "Convenience Stores", "Pharmacies and Drug Stores",
    "Cosmetics, Beauty Supplies, and Perfume Stores",
    "Depository Credit Intermediation",
    "Electronics and Appliance Stores",
    "Home Furnishings Stores", "Furniture Stores",
    "Automobile Dealers", "Tire Dealers",
    "Investigation and Security Services",
    "General Automotive Repair", "Car Washes",
    "Home Centers", "Paint and Wallpaper Stores",
    "Bus and Other Motor Vehicle Transit Systems",
    "Gasoline Stations with Convenience Stores", "Other Gasoline Stations",
    "Building Contractors",
    "Manufacturing", "Nature Parks and Other Similar Institutions",
    "Amusement and Theme Parks", "Golf Courses and Country Clubs",
    "Couriers and Express Delivery Services"),
  category_risk = c(
    1.8, 1.7, 1.5, 1.5, 1.5, 1.5, 1.5, 1.4, 1.4,
    1.4, 1.3, 1.3, 1.3, 1.2, 1.1, 1.2, 1.3, 1.2,
    1.2, 1.1, 1.2, 1.1, 1.2, 1.1, 1.1, 1.1, 1.1,
    1.1, 1.0, 1.1, 1.0, 1.0, 1.0, 1.0, 1.0, 1.0,
    1.0, 0.9, 0.9, 0.9, 0.8, 0.9, 0.9, 0.8, 0.8,
    0.8, 0.8, 0.8, 0.7, 0.7, 0.7, 0.7),
  enclosed_mod = c(
    1.0, 1.0, 1.0, 1.0, 1.0, 1.0, 1.0, 1.0, 1.0,
    1.0, 1.0, 1.0, 1.0, 1.0, 1.0, 1.0, 1.0, 1.0,
    1.0, 1.0, 1.0, 1.0, 1.0, 1.0, 1.0, 1.0, 1.0,
    1.0, 1.0, 1.0, 1.0, 1.0, 1.0, 1.0, 1.0, 1.0,
    0.0, 0.5, 1.0, 0.5, 0.5, 1.0, 1.0, 0.0, 0.5,
    0.5, 0.0, 1.0, 0.0, 0.5, 0.0, 0.0),
  effective_hours = c(
    8, 14, 12, 12, 16, 10, 10, 10, 10,
    9, 9, 9, 11, 10, 10, 8, 12, 10,
    16, 12, 11, 11, 10, 11, 11, 11, 11,
    10, 16, 24, 14, 11, 10, 11, 11, 11,
    11, 10, 10, 10, 12, 12, 11, 16, 24,
    24, 10, 10, 14, 10, 12, 10))

# Fallback: TOP_CATEGORY risk for POIs missing SUB_CATEGORY.
# Only category_risk is provided here; enclosed_mod / effective_hours fall
# back to defaults (1.0 and 12) inside process_pois().
top_category_risk <- data.table(
  top_category = c(
    "Religious Organizations", "Drinking Places (Alcoholic Beverages)",
    "Restaurants and Other Eating Places", "Fitness and Recreational Sports Centers",
    "Personal Care Services", "Offices of Dentists",
    "Offices of Physicians", "Offices of Other Health Practitioners",
    "Child Day Care Services", "Elementary and Secondary Schools",
    "Museums, Historical Sites", "Other Amusement and Recreation",
    "Limited-Service Restaurants", "Specialty Food Stores",
    "Other Schools and Instruction", "Clothing Stores",
    "Sporting Goods Stores", "Jewelry Stores",
    "Drycleaning and Laundry Services", "Grocery Stores", "Supermarkets",
    "Health and Personal Care Stores", "Depository Credit Intermediation",
    "Electronics and Appliance Stores", "Home Furnishings Stores",
    "Furniture Stores", "Office-based services", "Automobile Dealers",
    "Investigation and Security Services", "Automotive Repair and Maintenance",
    "Building Material Dealers", "Urban Transit Systems",
    "Gasoline Stations", "Building Contractors",
    "Manufacturing", "Nature Parks",
    "Amusement and Theme Parks", "Golf Courses and Country Clubs",
    "Couriers and Express Delivery"),
  category_risk_top = c(
    1.8, 1.7, 1.5, 1.5, 1.5, 1.4, 1.3, 1.3,
    1.3, 1.2, 1.2, 1.2, 1.2, 1.2, 1.2, 1.1,
    1.1, 1.1, 1.1, 1.0, 1.0, 1.0, 1.0, 1.0,
    1.0, 1.0, 1.0, 1.0, 0.9, 0.9, 0.9, 0.8,
    0.8, 0.8, 0.8, 0.7, 0.7, 0.7, 0.7))
```

# JSON Parsers

```{r parsers}
parse_json_safe <- function(x) tryCatch(fromJSON(x), error = function(e) NULL)

# Merge a vector of JSON dicts (key -> count) by summing values per key.
# Returns a single JSON string suitable for downstream parsers.
merge_count_jsons <- function(json_vec) {
  json_vec <- json_vec[!is.na(json_vec) & nchar(json_vec) > 2]
  if (length(json_vec) == 0) return(NA_character_)
  agg <- list()
  for (j in json_vec) {
    p <- parse_json_safe(j)
    if (is.null(p) || length(p) == 0) next
    nms <- names(p)
    if (is.null(nms)) next
    vals <- suppressWarnings(as.numeric(unlist(p)))
    keep <- !is.na(vals)
    if (!any(keep)) next
    for (i in which(keep)) {
      k <- nms[i]
      if (is.null(k) || is.na(k) || nchar(k) == 0) next
      agg[[k]] <- (if (is.null(agg[[k]])) 0 else agg[[k]]) + vals[i]
    }
  }
  if (length(agg) == 0) return(NA_character_)
  toJSON(agg, auto_unbox = TRUE)
}

# Sum element-wise a vector of JSON arrays of fixed length n.
# Returns a JSON array string of length n.
sum_array_jsons <- function(json_vec, n) {
  acc <- numeric(n)
  any_ok <- FALSE
  for (j in json_vec) {
    if (is.na(j) || nchar(j) < 3) next
    v <- suppressWarnings(as.numeric(parse_json_safe(j)))
    if (length(v) == 0 || all(is.na(v))) next
    v <- v[1:min(length(v), n)]
    v[is.na(v)] <- 0
    if (length(v) < n) v <- c(v, rep(0, n - length(v)))
    acc <- acc + v
    any_ok <- TRUE
  }
  if (!any_ok) return(NA_character_)
  toJSON(acc, auto_unbox = FALSE)
}

# Reduce a 168-element VISITS_BY_EACH_HOUR (7 days × 24 hours) JSON
# to a 24-element hour-of-day vector by summing across days.
# Returns JSON array of length 24.
collapse_vbeh_to_24 <- function(j) {
  if (is.na(j) || nchar(j) < 3) return(NA_character_)
  v <- suppressWarnings(as.numeric(parse_json_safe(j)))
  if (length(v) < 24) return(NA_character_)
  v[is.na(v)] <- 0
  if (length(v) >= 168) {
    m <- matrix(v[1:168], nrow = 7, byrow = TRUE)  # rows = days
    h24 <- colSums(m)
  } else {
    h24 <- v[1:24]
  }
  toJSON(h24, auto_unbox = FALSE)
}

parse_wd_frac <- function(j) {
  # Expects 7-element JSON array (Mon..Sun, ADVAN convention).
  v <- tryCatch(as.numeric(fromJSON(j)), error = function(e) NULL)
  if (is.null(v) || length(v) < 7 || sum(v, na.rm = TRUE) == 0) return(NA_real_)
  v <- v[1:7]; v[is.na(v)] <- 0
  tot <- sum(v); if (tot == 0) return(NA_real_)
  sum(v[1:5]) / tot
}

parse_hourly <- function(j) {
  h <- tryCatch(as.numeric(fromJSON(j)), error = function(e) NULL)
  if (is.null(h) || length(h) < 24 || sum(h, na.rm = TRUE) == 0)
    return(list(pk = NA_integer_, hc = NA_real_))
  h <- h[1:24]; h[is.na(h)] <- 0; props <- h / sum(h)
  list(pk = which.max(h) - 1L, hc = sum(props^2))
}

parse_daily_cv <- function(j) {
  v <- tryCatch(as.numeric(fromJSON(j)), error = function(e) NULL)
  if (is.null(v) || length(v) < 2 || sum(v, na.rm = TRUE) == 0)
    return(list(cv = NA_real_, wd_dom = NA_real_))
  v[is.na(v)] <- 0
  props <- v / sum(v)
  list(cv = sd(v) / mean(v), wd_dom = max(props))
}

parse_bucketed_dwell <- function(j) {
  d <- parse_json_safe(j)
  if (is.null(d) || length(d) == 0) return(NA_real_)
  midpoints <- c(2.5, 12.5, 40, 150, 360)  # <5, 5-20, 21-60, 61-240, >240 (min)
  counts <- suppressWarnings(as.numeric(unlist(d)))
  if (length(counts) > 5) counts <- counts[1:5]
  counts[is.na(counts)] <- 0
  if (sum(counts) > 0)
    sum(midpoints[1:length(counts)] * counts) / sum(counts)
  else NA_real_
}

# Parse DEVICE_TYPE JSON {"ios":N,"android":N,...} -> ios fraction
parse_device_ios_frac <- function(j) {
  if (is.na(j) || nchar(j) < 3) return(NA_real_)
  p <- parse_json_safe(j)
  if (is.null(p) || length(p) == 0) return(NA_real_)
  nms <- tolower(names(p))
  vals <- suppressWarnings(as.numeric(unlist(p)))
  vals[is.na(vals)] <- 0
  ios <- sum(vals[nms == "ios"])
  android <- sum(vals[nms == "android"])
  tot <- ios + android
  if (tot <= 0) return(NA_real_)
  ios / tot
}

# Long-format home-CBG visitor table from one merged JSON per POI.
parse_visitor_home_cbgs <- function(poi_home_dt, state) {
  poi_home_dt <- poi_home_dt[!is.na(visitor_home_cbgs_json) & nchar(visitor_home_cbgs_json) > 2]
  if (nrow(poi_home_dt) == 0) return(data.table())
  state_fips_2 <- substr(get_metro_fips(state, 5)[1], 1, 2)
  chunks <- split(seq_len(nrow(poi_home_dt)), ceiling(seq_len(nrow(poi_home_dt)) / 5000))
  rbindlist(lapply(chunks, function(idx) {
    out <- poi_home_dt[idx][, {
      parsed <- parse_json_safe(.BY$visitor_home_cbgs_json)
      if (is.null(parsed) || length(parsed) == 0)
        data.table(home_cbg = character(0), visitors = integer(0))
      else
        data.table(home_cbg = names(parsed),
                   visitors = as.integer(unlist(parsed)))
    }, by = .(poi_cbg, poi_lon, poi_lat, visits_total, median_dwell, brand,
              naics_code, visitor_home_cbgs_json)]
    out[, visitor_home_cbgs_json := NULL]
    out[!is.na(home_cbg) & substr(home_cbg, 1, 2) == state_fips_2 & visitors > 0]
  }))
}

parse_covisitation <- function(od_raw, json_col) {
  rows <- od_raw[!is.na(get(json_col)) & nchar(get(json_col)) > 2 & nchar(brand) > 0]
  if (nrow(rows) == 0) return(data.table())
  rbindlist(lapply(seq_len(nrow(rows)), function(i) {
    parsed <- parse_json_safe(rows[i][[json_col]])
    if (is.null(parsed) || length(parsed) == 0) return(NULL)
    data.table(source_brand = rows[i]$brand,
               target_brand = names(parsed),
               covisit_count = suppressWarnings(as.integer(unlist(parsed))))
  }))
}

parse_daytime_visitors <- function(j) {
  d <- parse_json_safe(j)
  if (is.null(d) || length(d) == 0) return(0L)
  v <- suppressWarnings(as.integer(unlist(d)))
  sum(v, na.rm = TRUE)
}
```

# Per-POI Wave Aggregation

```{r aggregate_helpers}
# Aggregate weekly rows for one POI into a single row covering the wave.
# Scalar fields summed/weighted; JSON fields merged via the helpers above.
aggregate_weekly_to_poi <- function(weekly_dt) {
  weekly_dt[, w := pmax(visits_total_w, 0)]  # weight by weekly visits
  weekly_dt[, ios_frac := vapply(device_type_json, parse_device_ios_frac,
                                 numeric(1))]
  agg <- weekly_dt[, .(
    poi_lon         = poi_lon[1],
    poi_lat         = poi_lat[1],
    top_category    = top_category[1],
    sub_category    = sub_category[1],
    brand           = brand[1],
    naics_code      = naics_code[1],
    poi_cbg         = poi_cbg[1],
    region          = region[1],

    visits_total    = sum(visits_total_w, na.rm = TRUE),
    visitors_unique = sum(visitors_unique_w, na.rm = TRUE),
    median_dwell    = if (sum(w, na.rm = TRUE) > 0)
                        weighted.mean(median_dwell_w, w, na.rm = TRUE)
                      else NA_real_,
    distance_from_home = if (sum(w, na.rm = TRUE) > 0)
                           weighted.mean(distance_from_home_w, w, na.rm = TRUE)
                         else NA_real_,
    ios_frac        = if (sum(w, na.rm = TRUE) > 0)
                        weighted.mean(ios_frac, w, na.rm = TRUE)
                      else NA_real_,

    # Merged JSONs (per-POI, wave-level)
    visits_by_day_json       = sum_array_jsons(visits_by_day_json_w, 7),
    popularity_by_hour_json  = {
      h24 <- vapply(visits_by_each_hour_json_w, collapse_vbeh_to_24,
                    character(1))
      sum_array_jsons(h24, 24)
    },
    visitor_home_cbgs_json   = merge_count_jsons(visitor_home_cbgs_json_w),
    visitor_daytime_cbgs_json = merge_count_jsons(visitor_daytime_cbgs_json_w),
    bucketed_dwell_times_json = merge_count_jsons(bucketed_dwell_times_json_w),
    related_same_day_brand_json  = merge_count_jsons(related_same_day_brand_json_w),
    related_same_week_brand_json = merge_count_jsons(related_same_week_brand_json_w)
  ), by = .(persistent_id)]

  # Derived weekday/weekend counts from merged VISITS_BY_DAY (Mon..Sun)
  agg[, wd_frac := vapply(visits_by_day_json, parse_wd_frac, numeric(1))]
  agg
}

lookup_parse <- function(dt, json_col, parse_fn, out_cols) {
  valid <- !is.na(dt[[json_col]]) & nchar(dt[[json_col]]) > 2
  if (!any(valid)) { dt[, (json_col) := NULL]; return(dt) }
  uniq <- unique(dt[[json_col]][valid])
  lk <- rbindlist(lapply(uniq, function(j) {
    res <- parse_fn(j)
    if (is.list(res)) c(setNames(list(j), json_col), res)
    else c(setNames(list(j), json_col), setNames(list(res), out_cols))
  }), fill = TRUE)
  dt <- merge(dt, lk, by = json_col, all.x = TRUE)
  dt[, (json_col) := NULL]
  dt
}

# Apply category risk, derived metrics, and wd/we splits to the per-POI table.
# `state` ("NY" or "UT") is required for the building-footprint area cache.
process_pois <- function(poi_dt, state) {
  # NAICS lookup: category_risk + enclosed_mod + effective_hours in one merge.
  poi_dt <- merge(poi_dt, naics_lookup, by = "sub_category", all.x = TRUE)
  # category_risk fallback via TOP_CATEGORY
  poi_dt <- merge(poi_dt, top_category_risk, by = "top_category", all.x = TRUE)
  poi_dt[is.na(category_risk),    category_risk    := category_risk_top]
  poi_dt[is.na(category_risk),    category_risk    := 1.0]
  poi_dt[is.na(enclosed_mod),     enclosed_mod     := 1.0]   # default indoor
  poi_dt[is.na(effective_hours),  effective_hours  := 12]    # default 12 h/day
  poi_dt[, category_risk_top := NULL]

  # Per-POI footprint area via MS Building Footprints (cached).
  area_dt <- build_poi_area_cache(poi_dt, state)
  poi_dt[area_dt, poi_area_sqm := i.poi_area_sqm, on = "persistent_id"]
  match_rate <- mean(!is.na(poi_dt$poi_area_sqm))
  message(sprintf("  POI footprint match rate: %.1f%% (%s/%s)",
                  100 * match_rate,
                  fmt(sum(!is.na(poi_dt$poi_area_sqm))),
                  fmt(nrow(poi_dt))))
  if (match_rate < 0.80)
    warning(sprintf("Low POI footprint coverage (%.1f%%); crowding factor will lean on median imputation.",
                    100 * match_rate))

  poi_dt[, dwell_hours := pmin(median_dwell / 60, 8)]
  poi_dt[is.na(dwell_hours) | dwell_hours <= 0, dwell_hours := 0.5]
  poi_dt[, activity_per_visit := (1 + dwell_hours) * category_risk * enclosed_mod]

  # No NORMALIZED_VISITS_BY_STATE_SCALING in Weekly+; visits_norm == raw.
  poi_dt[, visits_norm := as.numeric(visits_total)]
  poi_dt[is.na(visits_norm) | visits_norm <= 0, visits_norm := 0]
  poi_dt[, activity_total := visits_norm * activity_per_visit]
  poi_dt[is.na(distance_from_home) | distance_from_home < 0,
         distance_from_home := NA_real_]

  # Visit density (visits / m^2) and hourly contact density.
  poi_dt[, visit_density := fifelse(
    is.na(poi_area_sqm) | poi_area_sqm <= 0, NA_real_,
    visits_norm / poi_area_sqm)]
  poi_dt[, contact_density_hourly := fifelse(
    is.na(poi_area_sqm) | poi_area_sqm <= 0 | is.na(effective_hours) |
      effective_hours <= 0, NA_real_,
    visits_norm / poi_area_sqm / effective_hours)]

  # Weekday split
  poi_dt[is.na(wd_frac), wd_frac := 5/7]
  poi_dt[, visits_weekday := round(visits_norm * wd_frac)]
  poi_dt[, visits_weekend := visits_norm - visits_weekday][, wd_frac := NULL]

  # Hourly metrics from merged 24-hour JSON
  poi_dt[, c("peak_hour", "hourly_concentration") := .(NA_integer_, NA_real_)]
  poi_dt <- lookup_parse(poi_dt, "popularity_by_hour_json", parse_hourly,
                         c("pk", "hc"))
  poi_dt[!is.na(pk), peak_hour := pk]
  poi_dt[!is.na(hc), hourly_concentration := hc]
  poi_dt[, c("pk", "hc") := NULL]

  # Daily CV / dominance from merged 7-day JSON
  poi_dt[, c("daily_visit_cv", "weekday_dominance") := .(NA_real_, NA_real_)]
  poi_dt <- lookup_parse(poi_dt, "visits_by_day_json", parse_daily_cv,
                         c("cv", "wd_dom"))
  poi_dt[!is.na(cv), daily_visit_cv := cv][, cv := NULL]
  poi_dt[!is.na(wd_dom), weekday_dominance := wd_dom][, wd_dom := NULL]

  # Repeat ratio
  poi_dt[, repeat_ratio := visits_total / pmax(visitors_unique, 1L)]

  # Median dwell from merged BUCKETED_DWELL_TIMES if available (overrides
  # the visits-weighted MEDIAN_DWELL when present).
  poi_dt[, bdwell := vapply(bucketed_dwell_times_json,
                            parse_bucketed_dwell, numeric(1))]
  poi_dt[!is.na(bdwell), median_dwell := bdwell]
  poi_dt[, bdwell := NULL]

  # Device fractions (proportional split, not exclusive)
  poi_dt[is.na(ios_frac), ios_frac := 0.5]
  poi_dt[, visits_ios     := visits_norm * ios_frac]
  poi_dt[, visits_android := visits_norm * (1 - ios_frac)]
  poi_dt[, activity_ios     := activity_total * ios_frac]
  poi_dt[, activity_android := activity_total * (1 - ios_frac)]
  poi_dt[, activity_weekday := visits_weekday * activity_per_visit]
  poi_dt[, activity_weekend := visits_weekend * activity_per_visit]
  poi_dt
}

aggregate_to_grid <- function(dt) {
  dt[, .(
    n_pois = .N,
    visits_total       = sum(visits_total, na.rm = TRUE),
    visits_norm        = sum(visits_norm, na.rm = TRUE),
    visits_weekday     = sum(visits_weekday, na.rm = TRUE),
    visits_weekend     = sum(visits_weekend, na.rm = TRUE),
    activity_total     = sum(activity_total, na.rm = TRUE),
    activity_weekday   = sum(activity_weekday, na.rm = TRUE),
    activity_weekend   = sum(activity_weekend, na.rm = TRUE),
    visits_ios         = sum(visits_ios, na.rm = TRUE),
    visits_android     = sum(visits_android, na.rm = TRUE),
    activity_ios       = sum(activity_ios, na.rm = TRUE),
    activity_android   = sum(activity_android, na.rm = TRUE),
    avg_dwell                  = weighted.mean(dwell_hours, visits_norm, na.rm = TRUE),
    avg_category_risk          = weighted.mean(category_risk, visits_norm, na.rm = TRUE),
    enclosed_frac              = weighted.mean(enclosed_mod, visits_norm, na.rm = TRUE),
    avg_distance_from_home     = weighted.mean(distance_from_home, visits_norm, na.rm = TRUE),
    avg_poi_area               = mean(poi_area_sqm, na.rm = TRUE),
    avg_visit_density          = weighted.mean(visit_density, visits_norm, na.rm = TRUE),
    avg_peak_hour              = weighted.mean(peak_hour, visits_norm, na.rm = TRUE),
    avg_hourly_conc            = weighted.mean(hourly_concentration, visits_norm, na.rm = TRUE),
    avg_daily_cv               = weighted.mean(daily_visit_cv, visits_norm, na.rm = TRUE),
    avg_weekday_dominance      = weighted.mean(weekday_dominance, visits_norm, na.rm = TRUE),
    avg_repeat_ratio           = weighted.mean(repeat_ratio, visits_norm, na.rm = TRUE),
    avg_effective_hours        = weighted.mean(effective_hours, visits_norm, na.rm = TRUE),
    avg_contact_density_hourly = weighted.mean(contact_density_hourly, visits_norm, na.rm = TRUE)
  ), by = grid_id]
}
```

# Wave Driver

```{r process_wave}
# Pick weekly files whose Mon-Sun window overlaps the wave date range.
# Filename format: {YYYY-MM-DD}__dwllv2_{external_id}.parquet
select_files_for_range <- function(advan_files, wave_start, wave_end) {
  bn <- basename(advan_files)
  pk <- substr(bn, 1, 10)
  ok <- suppressWarnings(!is.na(as.Date(pk, format = "%Y-%m-%d")))
  pk <- as.Date(pk[ok], format = "%Y-%m-%d")
  files <- advan_files[ok]
  # Week covers [pk, pk+6]; keep if it intersects the wave window.
  keep <- (pk + 6) >= wave_start & pk <= wave_end
  files[keep]
}

process_wave <- function(wave_id, wave_months) {
  out_dir <- get_wave_dir(state, wave_id)
  dir.create(file.path(out_dir, "intermediate"), recursive = TRUE,
             showWarnings = FALSE)

  fips_5  <- get_metro_fips(state, 5)
  utm_crs <- get_utm_crs(state)
  wdr     <- get_wave_date_range(wave_months)
  total_n_days <- wdr$n_days

  grid_sf   <- st_read(file.path(CONFIG$grids_dir,
                                 sprintf("grid_100m_%s.gpkg", state)),
                       quiet = TRUE)
  grid_sf   <- st_transform(grid_sf, utm_crs)
  grid_join <- grid_sf[, "grid_id"]

  advan_files <- list.files(CONFIG$advan_dir, pattern = "\\.parquet$",
                            full.names = TRUE)
  wave_files <- select_files_for_range(advan_files, wdr$start, wdr$end)
  message(sprintf("[%s/%s] %d/%d weekly files in [%s, %s]",
                  state, wave_id, length(wave_files), length(advan_files),
                  wdr$start, wdr$end))
  if (length(wave_files) == 0)
    stop("No weekly files matched the wave window. Check filename format.")

  con <- dbConnect(duckdb(), dbdir = ":memory:")
  on.exit(dbDisconnect(con, shutdown = TRUE), add = TRUE)
  fsql <- paste0("'", fips_5, "'", collapse = ", ")
  region_filter <- if (state == "UT") "REGION = 'UT'"
                   else "REGION = 'NY'"

  ## ------------- Step 1: read all weekly POI rows for this wave -----------
  files_glob <- paste0("'", paste(wave_files, collapse = "','"), "'")
  weekly_query <- sprintf("
    SELECT
      CAST(PERSISTENT_ID AS VARCHAR)                    AS persistent_id,
      CAST(LONGITUDE AS DOUBLE)                         AS poi_lon,
      CAST(LATITUDE  AS DOUBLE)                         AS poi_lat,
      CAST(POI_CBG AS VARCHAR)                          AS poi_cbg,
      CAST(REGION AS VARCHAR)                           AS region,
      TOP_CATEGORY                                      AS top_category,
      COALESCE(SUB_CATEGORY,'')                         AS sub_category,
      COALESCE(BRAND,'')                                AS brand,
      COALESCE(NAICS_CODE,'')                           AS naics_code,

      DATE_RANGE_START                                  AS week_start,
      CAST(VISIT_COUNTS  AS DOUBLE)                     AS visits_total_w,
      CAST(VISITOR_COUNTS AS DOUBLE)                    AS visitors_unique_w,
      CAST(MEDIAN_DWELL  AS DOUBLE)                     AS median_dwell_w,
      CAST(DISTANCE_FROM_HOME AS DOUBLE)                AS distance_from_home_w,
      CAST(OPEN_DATE  AS VARCHAR)                       AS open_date,
      CAST(CLOSE_DATE AS VARCHAR)                       AS close_date,
      IS_DISTRIBUTOR                                    AS is_distributor,

      COALESCE(CAST(VISITS_BY_DAY        AS VARCHAR),'') AS visits_by_day_json_w,
      COALESCE(CAST(VISITS_BY_EACH_HOUR  AS VARCHAR),'') AS visits_by_each_hour_json_w,
      COALESCE(CAST(VISITOR_HOME_CBGS    AS VARCHAR),'') AS visitor_home_cbgs_json_w,
      COALESCE(CAST(VISITOR_DAYTIME_CBGS AS VARCHAR),'') AS visitor_daytime_cbgs_json_w,
      COALESCE(CAST(BUCKETED_DWELL_TIMES AS VARCHAR),'') AS bucketed_dwell_times_json_w,
      COALESCE(CAST(RELATED_SAME_DAY_BRAND  AS VARCHAR),'') AS related_same_day_brand_json_w,
      COALESCE(CAST(RELATED_SAME_WEEK_BRAND AS VARCHAR),'') AS related_same_week_brand_json_w,
      COALESCE(CAST(DEVICE_TYPE          AS VARCHAR),'') AS device_type_json
    FROM read_parquet([%s])
    WHERE %s
      AND SUBSTR(CAST(POI_CBG AS VARCHAR),1,5) IN (%s)
      AND VISIT_COUNTS > 0
      AND LONGITUDE IS NOT NULL AND LATITUDE IS NOT NULL
      AND PERSISTENT_ID IS NOT NULL",
    files_glob, region_filter, fsql)

  weekly_dt <- as.data.table(dbGetQuery(con, weekly_query))
  message(sprintf("  read %s weekly POI rows", fmt(nrow(weekly_dt))))
  if (nrow(weekly_dt) == 0) {
    return(list(state = state, wave_id = wave_id, n_pois = 0,
                n_days = total_n_days))
  }

  # Drop distribution-only locations (warehouses, depots): visit counts there
  # are dominated by trucks and don't reflect human contact density.
  n_before <- nrow(weekly_dt)
  weekly_dt <- weekly_dt[is.na(is_distributor) | is_distributor != 1]
  if (nrow(weekly_dt) < n_before)
    message(sprintf("  dropped %s distributor rows",
                    fmt(n_before - nrow(weekly_dt))))

  # Filter by per-POI active period: keep only weeks where the POI was open.
  # Defaults: OPEN_DATE = 1970-01-01 ("any time"), CLOSE_DATE = 2038-01-01.
  weekly_dt[, open_d  := suppressWarnings(as.Date(open_date))]
  weekly_dt[, close_d := suppressWarnings(as.Date(close_date))]
  weekly_dt[is.na(open_d),  open_d  := as.Date("1970-01-01")]
  weekly_dt[is.na(close_d), close_d := as.Date("2038-01-01")]
  weekly_dt[, week_end_d := as.Date(week_start) + 6L]
  n_before <- nrow(weekly_dt)
  weekly_dt <- weekly_dt[open_d <= week_end_d & close_d >= as.Date(week_start)]
  if (nrow(weekly_dt) < n_before)
    message(sprintf("  dropped %s rows outside POI active window",
                    fmt(n_before - nrow(weekly_dt))))
  weekly_dt[, c("open_d", "close_d", "week_end_d",
                "open_date", "close_date", "is_distributor") := NULL]
  if (nrow(weekly_dt) == 0) {
    return(list(state = state, wave_id = wave_id, n_pois = 0,
                n_days = total_n_days))
  }

  ## ------------- Step 2: build per-day daily_stats from VISITS_BY_DAY ------
  # Expand each weekly row's 7-vec to actual dates, keep only dates in wave.
  # Build the long table vectorized (no by-group) to avoid list-column issues.
  weekly_dt[, week_start_d := as.Date(week_start)]
  vbyday <- lapply(weekly_dt$visits_by_day_json_w, function(j) {
    v <- suppressWarnings(as.numeric(parse_json_safe(j)))
    if (length(v) >= 7) { v <- v[1:7]; v[is.na(v)] <- 0; v }
    else rep(0, 7)
  })
  vmat <- do.call(rbind, vbyday)  # n x 7
  n <- nrow(weekly_dt)
  start_int <- as.integer(wdr$start)
  end_int   <- as.integer(wdr$end)
  daily_dt <- data.table(
    day_int = rep(as.integer(weekly_dt$week_start_d), each = 7) +
              rep(0L:6L, times = n),
    vd      = as.numeric(t(vmat)),
    dwell   = rep(weekly_dt$median_dwell_w, each = 7)
  )
  daily_dt[, wt := vd]
  daily_dt <- daily_dt[day_int >= start_int & day_int <= end_int]

  ds_all <- daily_dt[, .(
    total_visits = sum(vd, na.rm = TRUE),
    median_dwell = if (sum(wt, na.rm = TRUE) > 0)
                     weighted.mean(dwell, wt, na.rm = TRUE) else NA_real_
  ), by = day_int]
  setorder(ds_all, day_int)
  ds_all[, date := as.Date(day_int, origin = "1970-01-01")]
  ds_all[, sim_day := day_int - start_int + 1L]
  ds_all[, source_month := format(date, "%Y-%m")]
  ds_all[, is_weekend := as.integer(lubridate::wday(date, week_start = 1) >= 6)]
  ds_all[, day_int := NULL]
  setcolorder(ds_all, c("sim_day", "date", "source_month", "is_weekend",
                        "total_visits", "median_dwell"))
  rm(daily_dt, vbyday, vmat); weekly_dt[, week_start_d := NULL]
  gc(verbose = FALSE)

  ## ------------- Step 2b: wave-level diurnal + dwell-bucket summaries ----
  # Sum the raw 168-vec VISITS_BY_EACH_HOUR across all weekly POI rows,
  # then write a 7x24 (dow x hour) diurnal table for EDA in 04.
  vbeh_acc <- numeric(168)
  for (j in weekly_dt$visits_by_each_hour_json_w) {
    if (is.na(j) || nchar(j) < 3) next
    v <- suppressWarnings(as.numeric(parse_json_safe(j)))
    if (length(v) < 168) next
    v <- v[1:168]; v[is.na(v)] <- 0
    vbeh_acc <- vbeh_acc + v
  }
  if (sum(vbeh_acc) > 0) {
    diurnal_dt <- data.table(
      dow    = rep(0:6, each = 24),    # Mon=0 ... Sun=6 (ADVAN convention)
      hour   = rep(0:23, times = 7),
      visits = vbeh_acc)
    fwrite(diurnal_dt, file.path(out_dir, "intermediate",
                                 "diurnal_profile.csv"))
  }

  # Sum BUCKETED_DWELL_TIMES bucket counts across all weekly POI rows.
  dwell_acc <- list()
  for (j in weekly_dt$bucketed_dwell_times_json_w) {
    if (is.na(j) || nchar(j) < 3) next
    p <- parse_json_safe(j)
    if (is.null(p) || length(p) == 0) next
    nms  <- names(p); if (is.null(nms)) next
    vals <- suppressWarnings(as.numeric(unlist(p)))
    keep <- !is.na(vals) & nzchar(nms)
    for (i in which(keep)) {
      k <- nms[i]
      dwell_acc[[k]] <- (if (is.null(dwell_acc[[k]])) 0 else dwell_acc[[k]]) + vals[i]
    }
  }
  if (length(dwell_acc) > 0) {
    dwell_dt <- data.table(bucket = names(dwell_acc),
                           count  = as.numeric(unlist(dwell_acc)))
    fwrite(dwell_dt, file.path(out_dir, "intermediate", "dwell_buckets.csv"))
  }

  ## ------------- Step 3: aggregate weekly rows to per-POI rows -----------
  poi_dt <- aggregate_weekly_to_poi(weekly_dt)
  rm(weekly_dt); gc(verbose = FALSE)
  message(sprintf("  %s unique POIs after wave aggregation", fmt(nrow(poi_dt))))

  poi_dt <- process_pois(poi_dt, state)

  ## ------------- Step 4: spatial join to grid + grid aggregation ---------
  poi_sf <- st_join(
    st_transform(st_as_sf(poi_dt[!is.na(poi_lon) & !is.na(poi_lat)],
                          coords = c("poi_lon", "poi_lat"), crs = 4326),
                 utm_crs),
    grid_join, left = FALSE)
  if (nrow(poi_sf) > 0) {
    grid_dt <- aggregate_to_grid(as.data.table(st_drop_geometry(poi_sf)))
    grid_dt[, `:=`(
      log_visits         = log1p(visits_norm),
      log_activity       = log1p(activity_total),
      log_ios_visits     = log1p(visits_ios),
      log_android_visits = log1p(visits_android),
      ios_frac           = visits_ios / pmax(visits_ios + visits_android, 1)
    )]
    grid_dt[, ses_contrast := ios_frac - 0.5]
    write_feather(grid_dt, file.path(out_dir, "intermediate",
                                     "grid_visits.feather"))
    message(sprintf("  wrote grid_visits.feather: %s grid cells",
                    fmt(nrow(grid_dt))))
  }

  fwrite(ds_all, file.path(out_dir, "intermediate", "daily_stats.csv"))
  message(sprintf("  wrote daily_stats.csv: %s days", fmt(nrow(ds_all))))

  ## ------------- Step 5: OD home-CBG pairs (per-POI, wave-merged) --------
  od_src <- poi_dt[!is.na(visitor_home_cbgs_json) &
                   nchar(visitor_home_cbgs_json) > 2,
                   .(poi_cbg, poi_lon, poi_lat, visits_total, median_dwell,
                     brand, naics_code, visitor_home_cbgs_json)]
  od_long <- parse_visitor_home_cbgs(od_src, state)
  if (nrow(od_long) > 0) {
    od_long[, `:=`(origin_fips = substr(home_cbg, 1, 5),
                   dest_fips   = substr(poi_cbg, 1, 5))]
    setnames(od_long, "visitors", "visitors")
    write_feather(od_long, file.path(out_dir, "intermediate",
                                     "od_home_cbg.feather"))
    message(sprintf("  wrote od_home_cbg.feather: %s pairs",
                    fmt(nrow(od_long))))
  }

  ## ------------- Step 6: brand co-visitation (day + week scope) ----------
  cov_src <- poi_dt[, .(
    brand = brand,
    covisit_json      = related_same_day_brand_json,
    covisit_week_json = related_same_week_brand_json
  )]
  cv_day  <- parse_covisitation(cov_src, "covisit_json")
  if (nrow(cv_day) > 0) cv_day[, scope := "day"]
  cv_week <- parse_covisitation(cov_src, "covisit_week_json")
  if (nrow(cv_week) > 0) cv_week[, scope := "week"]
  cv <- rbindlist(list(cv_day, cv_week), fill = TRUE)
  if (nrow(cv) > 0) {
    cv_agg <- cv[, .(covisit_count = sum(covisit_count), n_pois = .N),
                 by = .(source_brand, target_brand, scope)][order(-covisit_count)]
    write_feather(cv_agg, file.path(out_dir, "intermediate",
                                    "trip_chains.feather"))
    message(sprintf("  wrote trip_chains.feather: %s edges",
                    fmt(nrow(cv_agg))))
  }

  ## ------------- Step 7: daytime visitors per POI ------------------------
  dt_src <- poi_dt[!is.na(visitor_daytime_cbgs_json) &
                   nchar(visitor_daytime_cbgs_json) > 2]
  if (nrow(dt_src) > 0) {
    daytime_agg <- dt_src[, .(
      poi_lon = poi_lon[1], poi_lat = poi_lat[1],
      daytime_visitors = parse_daytime_visitors(visitor_daytime_cbgs_json)
    ), by = poi_cbg]
    write_feather(daytime_agg, file.path(out_dir, "intermediate",
                                         "daytime_visitors.feather"))
    message(sprintf("  wrote daytime_visitors.feather: %s POIs",
                    fmt(nrow(daytime_agg))))
  }

  list(state = state, wave_id = wave_id, n_pois = nrow(poi_dt),
       n_days = total_n_days)
}
```

# Run All

```{r run_all}
rbindlist(run_waves_by_state(process_wave), fill = TRUE)
```
