---
title: "04 - Mobility Data Diagnostics"
author: "Prince Oppong Boakye"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: show
    theme: flatly
---

```{r setup, include=FALSE}
source(here::here("helpers/_utils.R"))
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      fig.width = 14, fig.height = 8)
```

# Overview

Visualizes grid-level mobility outputs from 03 (Advan extraction).
No data transformations — purely diagnostic maps and distributions.

```{r init}
load_pipeline()
library(data.table); library(sf); library(arrow)
library(ggplot2); library(viridis); library(ggspatial); library(patchwork)

states <- unique(WAVE_RUNS$state)
spatial_data <- load_spatial(states)

# Pre-load wave-level grid_visits feathers (one per state x wave).
# Defensive dedupe by (wave_id, grid_id): some upstream feathers can carry
# duplicate grid_id rows, which would cause cartesian merges and overplotted maps.
visits_by_state <- setNames(lapply(states, function(state) {
  .this_state <- state
  st_waves <- WAVE_RUNS[which(WAVE_RUNS$state == .this_state), ]
  dt <- rbindlist(lapply(st_waves$wave_id, function(wid) {
    fp <- file.path(get_wave_dir(state, wid), "intermediate", "grid_visits.feather")
    if (!file.exists(fp)) return(NULL)
    d <- as.data.table(read_feather(fp))
    d[, wave_id := wid]
  }), fill = TRUE)
  n_dup <- nrow(dt) - nrow(unique(dt, by = c("wave_id", "grid_id")))
  if (n_dup > 0) {
    warning(sprintf("[%s] %d duplicate (wave_id, grid_id) rows in grid_visits; deduplicating (keeping first).",
                    state, n_dup))
    dt <- unique(dt, by = c("wave_id", "grid_id"))
  }
  dt
}), states)
```

## Maps (per wave)

```{r maps, fig.height=10, out.width="100%"}
for (state in states) {
  cbg_utm <- spatial_data[[state]]$cbg_utm
  grid_sf <- spatial_data[[state]]$grid_sf
  .this_state <- state
  st_waves <- WAVE_RUNS[which(WAVE_RUNS$state == .this_state), ]
  wave_ids <- unique(st_waves$wave_id)  # guard against duplicate WAVE_RUNS rows
  all_dt <- visits_by_state[[state]]
  make_mob_map <- function(plot_sf, fill_col, scale_fn, title_str) {
    ggplot() + cbg_base_layer(cbg_utm, state = state) +
      geom_sf(data = plot_sf, aes(fill = .data[[fill_col]]), color = NA) +
      scale_fn + map_extras() + labs(title = title_str) + theme_map()
  }

  for (wid in wave_ids) {
  gf <- merge(grid_sf[, c("grid_id", "fips_5")], all_dt[wave_id == wid], by = "grid_id", all.x = TRUE)
  gf$weekend_ratio <- gf$visits_weekend / pmax(gf$visits_weekday + gf$visits_weekend, 1)
  active <- gf[!is.na(gf$visits_total) & gf$visits_total > 0, ]

  if (state == "UT") {
    gf$log_visits   <- log1p(gf$visits_total)
    gf$log_activity <- log1p(gf$activity_total)
    gf$log_pois     <- log1p(gf$n_pois)
    cf1 <- ut_county_agg_sf(gf, "log_visits",   fun = mean)
    cf2 <- ut_county_agg_sf(gf, "log_activity",  fun = mean)
    cf3 <- ut_county_agg_sf(gf, "log_pois",      fun = mean)
    active$avg_dwell_   <- active$avg_dwell
    active$ios_frac_    <- active$ios_frac
    active$wknd_ratio_  <- active$weekend_ratio
    cf4 <- ut_county_agg_sf(active, "avg_dwell_",   fun = mean)
    cf5 <- ut_county_agg_sf(active, "ios_frac_",    fun = mean)
    cf6 <- ut_county_agg_sf(active, "wknd_ratio_",  fun = mean)

    p1 <- make_mob_map(cf1, "log_visits",   scale_fill_viridis_c(option="plasma", name="log(visits)"),   paste(wid, "- Visits (county mean)"))
    p2 <- make_mob_map(cf2, "log_activity", scale_fill_viridis_c(option="magma",  name="log(activity)"), paste(wid, "- Activity (county mean)"))
    p3 <- make_mob_map(cf3, "log_pois",     scale_fill_viridis_c(name="log(POIs)"),                      paste(wid, "- POI Density (county mean)"))
    p4 <- make_mob_map(cf4, "avg_dwell_",   scale_fill_viridis_c(option="inferno", name="Dwell (hrs)"),  paste(wid, "- Dwell Time (county mean)"))
    p5 <- make_mob_map(cf5, "ios_frac_",    scale_fill_gradient2(low="#d73027", mid="white", high="#4575b4", midpoint=0.5, name="iOS Frac"), paste(wid, "- SES Proxy"))
    p6 <- make_mob_map(cf6, "wknd_ratio_",  scale_fill_gradient2(low="#2166ac", mid="#f7f7f7", high="#b2182b", midpoint=2/7, name="Wknd Ratio"), paste(wid, "- Weekend Ratio"))
  } else {
    gf$log_visits   <- log1p(gf$visits_total)
    gf$log_activity <- log1p(gf$activity_total)
    gf$log_pois     <- log1p(gf$n_pois)
    p1 <- make_mob_map(gf,     "log_visits",    scale_fill_viridis_c(option="plasma",  name="log(visits)"),   paste(wid, "- Visits"))
    p2 <- make_mob_map(gf,     "log_activity",  scale_fill_viridis_c(option="magma",   name="log(activity)"), paste(wid, "- Activity"))
    p3 <- make_mob_map(gf,     "log_pois",      scale_fill_viridis_c(name="log(POIs)"),                       paste(wid, "- POI Density"))
    p4 <- make_mob_map(active, "avg_dwell",     scale_fill_viridis_c(option="inferno", name="Dwell (hrs)"),   paste(wid, "- Dwell Time"))
    p5 <- make_mob_map(active, "ios_frac",      scale_fill_gradient2(low="#d73027", mid="white", high="#4575b4", midpoint=0.5, name="iOS Frac"), paste(wid, "- SES Proxy (iOS)"))
    p6 <- make_mob_map(active, "weekend_ratio", scale_fill_gradient2(low="#2166ac", mid="#f7f7f7", high="#b2182b", midpoint=2/7, name="Wknd Ratio"), paste(wid, "- Weekend Ratio"))
  }
  print((p1 + p2) / (p3 + p4) / (p5 + p6))
  }
}
```

## Wave-on-Wave Change

Distribution of grid-level log2 visit ratios relative to the first wave (baseline).
Median (line) with IQR ribbon across grid cells, per wave. Above 0 = growth, below 0 = decline.
Useful for assessing whether the OD landscape stayed stable enough to pool kernel parameters across waves.

```{r change_trend, fig.width=10, fig.height=4}
trend_all <- rbindlist(lapply(states, function(state) {
  all_dt <- visits_by_state[[state]]
  waves  <- unique(all_dt$wave_id)
  if (length(waves) < 2) return(NULL)
  base_wid <- waves[1]
  base <- unique(all_dt[wave_id == base_wid, .(grid_id, v0 = visits_total)],
                 by = "grid_id")
  comp <- merge(all_dt, base, by = "grid_id")
  comp[, log2_ratio := log2((visits_total + 1) / (v0 + 1))]
  comp[, .(state = state,
           q25 = quantile(log2_ratio, 0.25, na.rm = TRUE),
           med = quantile(log2_ratio, 0.50, na.rm = TRUE),
           q75 = quantile(log2_ratio, 0.75, na.rm = TRUE)),
       by = wave_id]
}), fill = TRUE)

if (nrow(trend_all) > 0) {
  print(ggplot(trend_all, aes(x = wave_id, y = med, group = state, color = state)) +
    geom_ribbon(aes(ymin = q25, ymax = q75, fill = state),
                alpha = 0.15, color = NA) +
    geom_hline(yintercept = 0, linetype = "dashed", color = "grey40") +
    geom_line(linewidth = 0.7) + geom_point(size = 2) +
    labs(title = "Wave-on-wave mobility change (log2 ratio vs. baseline)",
         subtitle = "Line = grid-cell median; ribbon = IQR",
         x = "Wave", y = "log2((visits + 1) / (baseline + 1))") +
    theme_bw() + theme(legend.position = "bottom"))
}
```

## Distributions

```{r distributions, fig.width=14, fig.height=6}
for (state in states) {
  all_dt <- visits_by_state[[state]]
  dt <- all_dt[visits_total > 0]

  p1 <- ggplot(dt, aes(x = log1p(visits_total), fill = wave_id)) +
    geom_histogram(bins = 50, alpha = 0.6, position = "identity") +
    labs(title = paste(state, "— Log Visits by Wave"), x = "log(1 + visits)") + theme_bw()

  p2 <- ggplot(dt, aes(x = log1p(visits_total), y = log1p(activity_total))) +
    geom_point(alpha = 0.05, size = 0.3) + geom_smooth(method = "lm", color = "red") +
    facet_wrap(~wave_id) + labs(title = paste(state, "— Visits vs Activity")) + theme_bw()

  print(p1 + p2)
}
```

## OD Distance Distribution (visitor-weighted)

Per-pair home→POI distance, weighted by visitor count, faceted by wave. Red dashed line = visitor-weighted median.
Shifts across waves justify per-wave kernel σ, α refits in 06.

```{r od_distance, fig.width=14, fig.height=5}
for (state in states) {
  cbg_utm <- spatial_data[[state]]$cbg_utm
  cbg_centroid <- suppressWarnings(sf::st_centroid(cbg_utm))
  cbg_xy <- sf::st_coordinates(cbg_centroid)
  cbg_lookup <- data.table(
    home_cbg = if (state == "UT") cbg_centroid$GEOID10 else cbg_centroid$GEOID,
    home_x = cbg_xy[, 1], home_y = cbg_xy[, 2])

  .this_state <- state

  st_waves <- WAVE_RUNS[which(WAVE_RUNS$state == .this_state), ]
  od_all <- rbindlist(lapply(st_waves$wave_id, function(wid) {
    f <- file.path(get_wave_dir(state, wid), "intermediate", "od_home_cbg.feather")
    if (!file.exists(f)) return(NULL)
    dt <- as.data.table(read_feather(f)); dt[, wave_id := wid]
  }), fill = TRUE)
  if (nrow(od_all) == 0) next

  poi_unique <- unique(od_all[, .(poi_lon, poi_lat)])
  poi_proj <- sf::st_transform(
    sf::st_as_sf(poi_unique, coords = c("poi_lon","poi_lat"), crs = 4326),
    get_utm_crs(state))
  poi_xy <- sf::st_coordinates(poi_proj)
  poi_unique[, `:=`(poi_x = poi_xy[, 1], poi_y = poi_xy[, 2])]

  od_all <- merge(od_all, poi_unique, by = c("poi_lon","poi_lat"), all.x = TRUE)
  if (state == "UT") od_all[, home_key := substr(home_cbg, 1, 5)] else od_all[, home_key := home_cbg]
  setnames(cbg_lookup, "home_cbg", "home_key")
  od_all <- merge(od_all, cbg_lookup, by = "home_key", all.x = TRUE)
  od_all[, dist_km := sqrt((poi_x - home_x)^2 + (poi_y - home_y)^2) / 1000]
  od_all <- od_all[!is.na(dist_km) & dist_km > 0 & dist_km < 200]

  # Visitor-weighted median per wave (cumsum-based, no extra deps)
  med_by_wave <- od_all[order(dist_km),
    .(med = dist_km[which.min(abs(cumsum(visitors) / sum(visitors) - 0.5))]),
    by = wave_id]

  print(ggplot(od_all, aes(x = dist_km, weight = visitors)) +
    geom_histogram(bins = 50, fill = "steelblue", alpha = 0.75) +
    geom_vline(data = med_by_wave, aes(xintercept = med),
               color = "red", linetype = "dashed", inherit.aes = FALSE) +
    geom_text(data = med_by_wave,
              aes(x = med, y = Inf, label = sprintf("med=%.1f km", med)),
              vjust = 1.5, hjust = -0.05, color = "red", size = 3,
              inherit.aes = FALSE) +
    facet_wrap(~ wave_id, scales = "free_y") +
    scale_x_log10() +
    labs(title = paste(state, "— OD Distance (visitor-weighted)"),
         x = "Distance home → POI (km, log10)", y = "Visitor count") +
    theme_bw())
}
```

## OD Flow Map (flowmapblue)

Home → POI flows rendered with [flowmap.blue](https://flowmap.blue) via the
`flowmapblue` R wrapper. Aggregated **region × region** (CBG×CBG for NY,
county×county for UT) — i.e. a true OD matrix, not point-to-point — following
the LOOKUP/04 pattern. One widget per (state, wave); top-1500 strongest flows.
Skipped if `flowmapblue` not installed.

```{r od_flowmap, eval=requireNamespace("flowmapblue", quietly=TRUE)}
# NOTE: hard-coded for local runs; strip before pushing to remote.
MAPBOX_TOKEN <- "pk.eyJ1IjoicHJpbmNlb2IiLCJhIjoiY21tcmp6N210MHNnZTJvc2llN3l1ZWRobSJ9.aE_ZlsJYZVPHPmCaQtM3VA"

for (state in states) {
  # Build a single region->lon/lat lookup from the spatial layer.
  # NY: CBG (12-digit GEOID); UT: county (5-digit FIPS_STR / GEOID10).
  reg_sf <- spatial_data[[state]]$cbg_utm
  reg_id_col <- if (state == "UT") "GEOID10" else "GEOID"
  reg_ll <- suppressWarnings(sf::st_centroid(sf::st_transform(reg_sf, 4326)))
  reg_xy <- sf::st_coordinates(reg_ll)
  reg_lookup <- data.table(
    reg_id = as.character(reg_ll[[reg_id_col]]),
    lon    = reg_xy[, 1],
    lat    = reg_xy[, 2])
  reg_lookup <- reg_lookup[!is.na(reg_id)]
  reg_lookup <- unique(reg_lookup, by = "reg_id")

  # Region key: CBG keeps full 12-digit; county is 5-digit (truncate 12-digit
  # CBG inputs from od_home_cbg.feather to first 5 chars to align with county
  # FIPS in reg_lookup).
  to_key <- if (state == "UT") function(x) substr(x, 1, 5) else identity

  .this_state <- state

  st_waves <- WAVE_RUNS[which(WAVE_RUNS$state == .this_state), ]
  for (wid in st_waves$wave_id) {
    f <- file.path(get_wave_dir(state, wid), "intermediate", "od_home_cbg.feather")
    if (!file.exists(f)) next
    od <- as.data.table(read_feather(f))
    if (!nrow(od)) next

    # Aggregate visitors at region x region (origin home -> dest CBG/county)
    od[, `:=`(origin = to_key(home_cbg), dest = to_key(poi_cbg))]
    flows <- od[!is.na(origin) & !is.na(dest) & origin != "" & dest != "" &
                  visitors > 0,
                .(count = sum(visitors, na.rm = TRUE)),
                by = .(origin, dest)]
    if (!nrow(flows)) next

    # Top-1500 strongest pairs (keeps widget responsive for NY's ~6k CBGs)
    setorder(flows, -count)
    flows <- head(flows, 1500)

    # Build locations: only ids actually used in flows
    used_ids <- unique(c(flows$origin, flows$dest))
    locs <- reg_lookup[reg_id %in% used_ids,
                       .(id = reg_id, name = reg_id, lon, lat)]
    if (!nrow(locs)) next
    # Drop flows whose endpoints are missing centroids
    flows <- flows[origin %in% locs$id & dest %in% locs$id]
    if (!nrow(flows)) next

    # Persist OD pairs + locations alongside the widget for downstream reuse.
    out_dir <- file.path(get_wave_dir(state, wid), "intermediate")
    fwrite(flows, file.path(out_dir, "od_flow_pairs.csv"))
    fwrite(locs,  file.path(out_dir, "od_flow_locations.csv"))

    cat(sprintf("\n\n### %s \u2014 %s (%s flows, %s nodes)\n\n",
                state, mlabels[wid], fmt(nrow(flows)), fmt(nrow(locs))))
    print(htmltools::tagList(flowmapblue::flowmapblue(
      locations         = as.data.frame(locs),
      flows             = as.data.frame(flows),
      mapboxAccessToken = MAPBOX_TOKEN,
      darkMode          = TRUE,
      clustering        = TRUE)))
  }
}
```

## Brand Co-visitation Network

Top brand→brand co-visitation edges (same-day scope) per wave, with Louvain communities.
Honest framing: these are unordered co-visit counts (top-20 brands per POI), **not** ordered trip chains.
Communities reveal *trip-purpose bundles* — e.g., essentials (grocery + pharmacy) vs. discretionary (dining + retail).

```{r covisit_network, fig.width=12, fig.height=8}
library(igraph)
top_n_edges <- 80
brand_communities <- list()  # E10: collect Louvain memberships per (state, wave)
for (state in states) {
  .this_state <- state
  st_waves <- WAVE_RUNS[which(WAVE_RUNS$state == .this_state), ]
  pick_waves <- unique(st_waves$wave_id)
  for (wid in pick_waves) {
    f <- file.path(get_wave_dir(state, wid), "intermediate", "trip_chains.feather")
    if (!file.exists(f)) next
    edges <- as.data.table(read_feather(f))
    edges <- edges[scope == "day" &
                   !is.na(source_brand) & !is.na(target_brand) &
                   source_brand != "" & target_brand != "" &
                   source_brand != target_brand]
    if (nrow(edges) == 0) next

    # undirected: collapse symmetric pairs
    edges[, c("a","b") := .(pmin(source_brand, target_brand),
                            pmax(source_brand, target_brand))]
    edges <- edges[, .(weight = sum(covisit_count)), by = .(a, b)]
    setorder(edges, -weight)
    edges <- head(edges, top_n_edges)

    g <- graph_from_data_frame(edges[, .(a, b, weight)], directed = FALSE)
    cl <- cluster_louvain(g, weights = E(g)$weight)
    V(g)$community <- membership(cl)
    V(g)$degree    <- strength(g, weights = E(g)$weight)

    # E10: persist Louvain membership for cross-wave ARI/NMI.
    brand_communities[[paste(state, wid, sep = "_")]] <- data.table(
      state = state, wave_id = wid,
      brand = V(g)$name, community = as.integer(V(g)$community))

    pal <- viridis::viridis(max(V(g)$community))
    par(mar = c(0,0,2,0))
    set.seed(1)
    plot(g,
         layout         = layout_with_fr(g, weights = E(g)$weight),
         vertex.color   = pal[V(g)$community],
         vertex.size    = 4 + 8 * sqrt(V(g)$degree / max(V(g)$degree)),
         vertex.label.cex = 0.7,
         vertex.label.color = "black",
         vertex.frame.color = NA,
         edge.width     = 0.3 + 3 * (E(g)$weight / max(E(g)$weight)),
         edge.color     = adjustcolor("grey50", alpha = 0.4),
         main = sprintf("%s %s — Brand co-visitation (top %d edges, %d communities)",
                        state, wid, top_n_edges, length(unique(V(g)$community))))
  }
}

# E10: Cross-wave ARI/NMI of Louvain memberships on the brand intersection.
if (length(brand_communities) >= 2) {
  bc_all <- rbindlist(brand_communities, fill = TRUE)
  fwrite(bc_all, file.path("outputs", "spatiotemporal_risk",
                            "brand_louvain_memberships.csv"))
  ari_rows <- list()
  for (st in unique(bc_all$state)) {
    bc_st <- bc_all[state == st]
    waves_st <- sort(unique(bc_st$wave_id))
    if (length(waves_st) < 2) next
    for (i in seq_len(length(waves_st) - 1L)) {
      w1 <- waves_st[i]; w2 <- waves_st[i + 1L]
      m1 <- bc_st[wave_id == w1]
      m2 <- bc_st[wave_id == w2]
      common <- intersect(m1$brand, m2$brand)
      if (length(common) < 5) next
      c1 <- m1[match(common, brand), community]
      c2 <- m2[match(common, brand), community]
      ari <- igraph::compare(c1, c2, method = "adjusted.rand")
      nmi <- igraph::compare(c1, c2, method = "nmi")
      ari_rows[[length(ari_rows) + 1]] <- data.table(
        state = st, wave_a = w1, wave_b = w2,
        n_brands_common = length(common),
        ARI = ari, NMI = nmi)
    }
  }
  if (length(ari_rows) > 0) {
    ari_dt <- rbindlist(ari_rows, fill = TRUE)
    fwrite(ari_dt, file.path("outputs", "spatiotemporal_risk",
                              "brand_louvain_ari.csv"))
    print(knitr::kable(ari_dt, format = "pipe", digits = 3,
                caption = "Cross-wave Louvain stability (ARI / NMI on brand intersection)"))
  }
}
```

## Daily Visits Time Series

Single black line per wave window; weekends shaded pink (Sat 00:00 → Mon 00:00).
Style follows the rect-shaded approach from *Visualizing Time Series Data with
ggplot2* — minimalistic, lets the weekday/weekend rhythm read at a glance
without color/shape clutter.

```{r daily_timeseries, fig.width=10, fig.height=3.5}
for (state in states) {
  .this_state <- state
  st_waves <- WAVE_RUNS[which(WAVE_RUNS$state == .this_state), ]
  for (wid in st_waves$wave_id) {
    f <- file.path(get_wave_dir(state, wid), "intermediate", "daily_stats.csv")
    if (!file.exists(f)) next
    ds <- fread(f); ds[, date := as.Date(date)]
    if (!nrow(ds)) next
    wknd <- ds[wday(date) == 7, .(xmin = date, xmax = date + 2)]

    print(ggplot(ds, aes(x = date, y = total_visits)) +
      geom_rect(data = wknd, inherit.aes = FALSE,
                aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf),
                fill = "grey85", alpha = 0.45) +
      geom_line(linewidth = 0.5, color = "grey20") +
      scale_y_continuous(labels = scales::comma) +
      scale_x_date(date_labels = "%b %d", date_breaks = "2 weeks", expand = c(0, 0)) +
      labs(title = paste(state, "—", mlabels[wid]),
           subtitle = "Daily visit volume (weekends shaded)",
           x = NULL, y = "Total Visits") +
      theme_minimal(base_size = 11) +
      theme(plot.title = element_text(face = "bold"),
            panel.grid.minor = element_blank(),
            axis.text.x = element_text(angle = 30, hjust = 1)))
  }
}
```

## Diurnal Visit Profile (day-of-week × hour)

Visit-weighted heatmap of visits per (day-of-week, hour-of-day), one panel per
wave. Sourced from `intermediate/diurnal_profile.csv` (summed `VISITS_BY_EACH_HOUR`
across all weekly POI rows). Highlights how nightlife/curfew, work-from-home,
and weekend patterns shifted across pandemic phases. Skipped silently if
03 has not been re-run with the new summary writer.

```{r diurnal_heatmap, fig.width=10, fig.height=3.5}
dow_lab <- c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun")
for (state in states) {
  .this_state <- state
  st_waves <- WAVE_RUNS[which(WAVE_RUNS$state == .this_state), ]
  for (wid in st_waves$wave_id) {
    f <- file.path(get_wave_dir(state, wid), "intermediate",
                   "diurnal_profile.csv")
    if (!file.exists(f)) next
    dp <- fread(f); if (!nrow(dp) || sum(dp$visits) == 0) next
    dp[, dow_f := factor(dow_lab[dow + 1L], levels = dow_lab)]
    dp[, share := visits / sum(visits)]
    print(ggplot(dp, aes(x = hour, y = dow_f, fill = share)) +
      geom_tile() +
      scale_fill_viridis_c(option = "magma", labels = scales::percent_format(0.1)) +
      scale_x_continuous(breaks = seq(0, 23, 3), expand = c(0, 0)) +
      scale_y_discrete(limits = rev(dow_lab)) +
      labs(title = paste(state, "\u2014", mlabels[wid]),
           subtitle = "Share of weekly visits by day-of-week and hour",
           x = "Hour of day", y = NULL, fill = "Share") +
      theme_minimal(base_size = 11) +
      theme(plot.title = element_text(face = "bold"),
            panel.grid = element_blank()))
  }
}
```

## Dwell-Time Bucket Distribution

Wave-level breakdown of visit dwell-time buckets (`<5`, `5–20`, `21–60`,
`61–240`, `>240` minutes) summed across all POIs. Stacked-share bars compare
how exposure dose shifted across pandemic phases per state. Sourced from
`intermediate/dwell_buckets.csv` (re-run 03 to populate).

```{r dwell_buckets, fig.width=10, fig.height=4}
bucket_levels <- c("<5", "5-20", "21-60", "61-240", ">240")
# ADVAN keys: "<5", "5-20", "21-60", "61-240", ">240"
norm_bucket <- function(x) {
  x <- gsub("^<\\s*", "<", x)
  x <- gsub("^>\\s*", ">", x)
  x <- gsub("\\s", "", x)
  x
}
for (state in states) {
  .this_state <- state
  st_waves <- WAVE_RUNS[which(WAVE_RUNS$state == .this_state), ]
  rows <- rbindlist(lapply(st_waves$wave_id, function(wid) {
    f <- file.path(get_wave_dir(state, wid), "intermediate",
                   "dwell_buckets.csv")
    if (!file.exists(f)) return(NULL)
    d <- fread(f); if (!nrow(d)) return(NULL)
    d[, wave_id := wid]
  }), fill = TRUE)
  if (!nrow(rows)) next
  rows[, bucket := norm_bucket(bucket)]
  rows <- rows[bucket %in% bucket_levels]
  rows[, bucket := factor(bucket, levels = bucket_levels)]
  rows[, share  := count / sum(count), by = wave_id]
  # Build wave_label factor with unique, ordered levels (defensive against
  # duplicate labels or repeated wave_ids in st_waves).
  wave_order  <- unique(as.character(st_waves$wave_id))
  level_lbls  <- unique(unname(mlabels[wave_order]))
  rows[, wave_label := factor(unname(mlabels[wave_id]), levels = level_lbls)]
  print(ggplot(rows, aes(x = wave_label, y = share, fill = bucket)) +
    geom_col(width = 0.7) +
    scale_y_continuous(labels = scales::percent_format(1), expand = c(0, 0)) +
    scale_fill_brewer(palette = "YlOrRd", direction = 1) +
    labs(title = paste(state, "\u2014 Dwell-time bucket shares per wave"),
         subtitle = "Share of visits by dwell-time bucket (minutes)",
         x = NULL, y = "Share of visits", fill = "Dwell (min)") +
    theme_minimal(base_size = 11) +
    theme(plot.title = element_text(face = "bold"),
          panel.grid.major.x = element_blank(),
          legend.position = "right"))
}
```

## Wave Summary

```{r summary_table}
for (state in states) {
  all_dt <- visits_by_state[[state]]
  print(all_dt[, .(
    active_cells = sum(visits_total > 0), total_visits = sum(visits_total),
    total_pois = sum(n_pois),
    med_dwell = round(median(avg_dwell[avg_dwell > 0], na.rm = TRUE), 2),
    med_ios   = round(median(ios_frac[ios_frac > 0], na.rm = TRUE), 3),
    pct_wknd  = round(100 * sum(visits_weekend) / pmax(sum(visits_total), 1), 1)
  ), by = wave_id])
}
```

## Mobility–Case Coupling (EDA)

Pre-fit exploratory diagnostics linking observed mobility to observed case
incidence. Style borrowed from Niraula's London EDA (spaghetti + bold mean
overlay, wave-boundary annotations); aggregations adapted to NY (MODZCTA) and
UT (county) supports.

```{r coupling_load}
# Per-state daily (geo_id x date) caserate, plus daily state-mean visits
# from daily_stats.csv. Builds: case_long[state], visits_long[state].
case_long <- list(); visits_long <- list()

for (state in states) {
  .this_state <- state
  st_waves <- WAVE_RUNS[which(WAVE_RUNS$state == .this_state), ]
  cl <- rbindlist(lapply(seq_len(nrow(st_waves)), function(i) {
    wid <- st_waves$wave_id[i]
    f <- file.path(get_wave_dir(state, wid), "tables", "case_rates.csv")
    if (!file.exists(f)) return(NULL)
    d <- fread(f)
    d[, `:=`(geo_id = as.character(geo_id), date = as.Date(date), wave_id = wid)]
    d
  }), fill = TRUE)
  if (nrow(cl)) cl <- cl[!is.na(case_rate)]
  case_long[[state]] <- cl

  vl <- rbindlist(lapply(st_waves$wave_id, function(wid) {
    f <- file.path(get_wave_dir(state, wid), "intermediate", "daily_stats.csv")
    if (!file.exists(f)) return(NULL)
    d <- fread(f)
    d[, `:=`(date = as.Date(date), wave_id = wid)]
    d
  }), fill = TRUE)
  visits_long[[state]] <- vl
}
```

```{r coupling_caserate_spaghetti, fig.width=10, fig.height=3.5}
# One plot per (state, wave): per-geo_id caserate spaghetti + state-mean overlay.
for (state in states) {
  cl <- case_long[[state]]
  if (!nrow(cl)) next
  for (wid in unique(cl$wave_id)) {
    clw <- cl[wave_id == wid]
    if (!nrow(clw)) next
    state_mean <- clw[, .(mean_cr = mean(case_rate, na.rm = TRUE)), by = date]
    p <- ggplot(clw, aes(x = date, y = case_rate, group = geo_id)) +
      geom_line(linewidth = 0.15, alpha = 0.35, color = "#1f78b4") +
      geom_line(data = state_mean, aes(x = date, y = mean_cr, group = 1),
                color = "black", linewidth = 0.9, linetype = "dashed",
                inherit.aes = FALSE) +
      scale_x_date(date_breaks = "2 weeks", date_labels = "%b %d") +
      labs(title = paste(state, "—", mlabels[wid]),
           subtitle = "Daily case rate per geo (spaghetti) with state mean overlay",
           x = NULL, y = "Cases / 100k") +
      theme_minimal(base_size = 11) +
      theme(plot.title = element_text(face = "bold"),
            panel.grid.minor = element_blank(),
            axis.text.x = element_text(angle = 30, hjust = 1))
    print(p)
  }
}
```

```{r coupling_visits_vs_cases, fig.width=10, fig.height=3.5}
# One plot per (state, wave): state-mean visits vs caserate, z-scored within
# the wave window; 14-day rolling Pearson correlation overlaid on a secondary
# y-axis.
zs <- function(x) (x - mean(x, na.rm = TRUE)) / sd(x, na.rm = TRUE)
roll_cor <- function(x, y, w = 14) {
  n <- length(x); out <- rep(NA_real_, n)
  for (i in seq.int(w, n)) {
    xi <- x[(i - w + 1):i]; yi <- y[(i - w + 1):i]
    if (sum(!is.na(xi) & !is.na(yi)) >= w / 2)
      out[i] <- suppressWarnings(cor(xi, yi, use = "pairwise.complete.obs"))
  }
  out
}
for (state in states) {
  cl <- case_long[[state]]; vl <- visits_long[[state]]
  if (!nrow(cl) || !nrow(vl)) next
  for (wid in unique(cl$wave_id)) {
    daily <- merge(
      cl[wave_id == wid, .(mean_cr = mean(case_rate, na.rm = TRUE)), by = date],
      vl[wave_id == wid, .(total_visits = sum(total_visits, na.rm = TRUE)), by = date],
      by = "date", all = TRUE)
    if (!nrow(daily)) next
    setorder(daily, date)
    daily[, `:=`(z_cr = zs(mean_cr), z_v = zs(total_visits))]
    daily[, rcor := roll_cor(mean_cr, total_visits, w = 14)]

    long <- melt(daily[, .(date, `Cases (z)` = z_cr, `Visits (z)` = z_v)],
                 id.vars = "date", variable.name = "series", value.name = "value")
    p <- ggplot(long, aes(x = date, y = value, color = series)) +
      geom_line(linewidth = 0.6, alpha = 0.9) +
      geom_line(data = daily, aes(x = date, y = rcor * 2),
                color = "grey30", linewidth = 0.4, linetype = "dotted",
                inherit.aes = FALSE) +
      geom_hline(yintercept = 0, color = "grey70", linewidth = 0.3) +
      scale_color_manual(values = c("Cases (z)" = "#b2182b", "Visits (z)" = "#1f78b4")) +
      scale_y_continuous(name = "Standardized value",
                         sec.axis = sec_axis(~ . / 2, name = "14-day rolling cor")) +
      scale_x_date(date_breaks = "2 weeks", date_labels = "%b %d") +
      labs(title = paste(state, "—", mlabels[wid]),
           subtitle = "State-mean cases vs visits (z-scored) and rolling correlation",
           x = NULL, color = NULL) +
      theme_minimal(base_size = 11) +
      theme(plot.title = element_text(face = "bold"),
            legend.position = "top",
            panel.grid.minor = element_blank(),
            axis.text.x = element_text(angle = 30, hjust = 1),
            axis.title.y.right = element_text(color = "grey30"))
    print(p)
  }
}
```

```{r coupling_cumulative_choropleth, fig.width=14, fig.height=5.5}
# Per-wave cumulative case rate, faceted choropleth. Uses MODZCTA (NY) or
# county (UT) shapefiles already referenced by load_spatial / CONFIG.
modzcta_sf <- st_read(file.path("data", "MODZCTA",
  "geo_export_5392c927-ee1e-4b9c-9787-2cad9456c83b.shp"), quiet = TRUE)

for (state in states) {
  cl <- case_long[[state]]
  if (!nrow(cl)) next
  # Sum of daily per-100k rates across the wave window (proxy for cumulative
  # incidence; exact only if all geos share the same n_days, which is the case
  # within a wave by construction).
  wave_cum <- cl[, .(cum_cr = sum(case_rate, na.rm = TRUE)),
                 by = .(geo_id, wave_id)]
  wave_cum[, wave_label := mlabels[wave_id]]
  if (state == "NY") {
    sf_obj <- modzcta_sf
    sf_obj$geo_id <- as.character(sf_obj$modzcta)
  } else if (state == "UT") {
    sf_obj <- spatial_data[[state]]$cbg_utm
    sf_obj$geo_id <- as.character(sf_obj$FIPS_STR)
  } else {
    next
  }
  joined <- merge(sf_obj, wave_cum, by = "geo_id", all.x = TRUE)
  p <- ggplot(joined) +
    geom_sf(aes(fill = cum_cr), color = NA) +
    facet_wrap(~ wave_label, nrow = 1) +
    scale_fill_viridis_c(option = "H", trans = "sqrt", na.value = "grey90",
                         labels = scales::label_number(accuracy = 1)) +
    labs(title = paste(state, "— Sum of daily case rates by wave (per 100k)"),
         fill = NULL) +
    theme_map() +
    theme(legend.position = "bottom",
          legend.key.width = grid::unit(2, "cm"),
          legend.key.height = grid::unit(3, "mm"))
  print(p)
}
```

```{r coupling_geo_bivariate, fig.width=14, fig.height=4.5}
# Cross-sectional: total visits vs mean caserate at the geo unit, faceted by
# wave. Builds grid -> geo_id once per state from the grid gpkg.
for (state in states) {
  vbs  <- visits_by_state[[state]]
  cl   <- case_long[[state]]
  if (!nrow(vbs) || !nrow(cl)) next

  grid_path <- file.path(CONFIG$grids_dir, sprintf("grid_100m_%s.gpkg", state))
  grid_geo  <- as.data.table(st_drop_geometry(st_read(grid_path, quiet = TRUE)))
  if (state == "NY") {
    xwalk <- fread(CONFIG$nyc_crosswalk_path)
    xwalk[, `:=`(cbg_geoid = as.character(cbg_geoid),
                 modzcta   = as.character(modzcta))]
    xwalk <- xwalk[!duplicated(cbg_geoid)]
    grid_geo[, cbg_geoid := as.character(cbg_geoid)]
    grid_geo <- merge(grid_geo, xwalk[, .(cbg_geoid, modzcta)],
                      by = "cbg_geoid", all.x = TRUE)
    grid_geo[, geo_id := modzcta]
  } else {
    grid_geo[, geo_id := as.character(fips_5)]
  }
  grid_geo <- grid_geo[!is.na(geo_id) & geo_id != "unknown" & geo_id != "36NA"]

  vbs2 <- merge(vbs, grid_geo[, .(grid_id, geo_id)], by = "grid_id", all.x = TRUE)
  vbs_geo <- vbs2[!is.na(geo_id),
                  .(geo_visits = sum(visits_total, na.rm = TRUE)),
                  by = .(wave_id, geo_id)]
  cr_geo <- cl[, .(geo_caserate = mean(case_rate, na.rm = TRUE)),
               by = .(wave_id, geo_id)]
  bv <- merge(vbs_geo, cr_geo, by = c("wave_id", "geo_id"))
  bv <- bv[geo_visits > 0 & geo_caserate > 0]
  bv[, wave_label := mlabels[wave_id]]

  ann <- bv[, .(
    r   = cor(log10(geo_visits), geo_caserate, use = "complete.obs"),
    n   = .N
  ), by = wave_label]
  ann[, lab := sprintf("r = %.2f  (n = %d)", r, n)]

  p <- ggplot(bv, aes(x = geo_visits, y = geo_caserate)) +
    geom_point(alpha = 0.5, size = 1.2, color = "#1f78b4") +
    geom_smooth(method = "lm", se = FALSE, color = "#b2182b", linewidth = 0.6) +
    geom_text(data = ann, aes(label = lab),
              x = -Inf, y = Inf, hjust = -0.1, vjust = 1.5,
              size = 3.4, color = "grey20", inherit.aes = FALSE) +
    facet_wrap(~ wave_label, nrow = 1) +
    scale_x_log10(labels = scales::label_number(scale_cut = scales::cut_short_scale())) +
    labs(title = paste(state, "— Total visits vs mean caserate at geo level (per wave)"),
         x = "Total wave visits (geo unit, log10)",
         y = "Mean wave caserate (per 100k)") +
    theme_bw() +
    theme(plot.title = element_text(face = "bold"),
          strip.background = element_rect(fill = "grey90"))
  print(p)
}
```

## Cross-Correlation Lag Scan (visits → cases)

```{r coupling_ccf_lagscan, fig.width=11, fig.height=5}
# AOAS Slater 2025 §3.2: scan lags ℓ ∈ [-3, +21] days for cor(visits_t,
# cases_{t+ℓ}); peak lag and its CI motivate the choice of mob_lag3/5/7.
# Per (state, wave): detrend both series with a centered 7-day rolling mean,
# then compute Pearson cor across lags. Persists the table (CSV) so 08a can
# read peak lags later if we want a data-driven `mob_lag_best` covariate.
detrend <- function(x, w = 7) {
  trend <- zoo::rollmean(x, k = w, fill = NA, align = "center")
  x - trend
}

ccf_rows <- list()
LAGS <- -3:28

for (state in states) {
  cl <- case_long[[state]]; vl <- visits_long[[state]]
  if (is.null(cl) || is.null(vl) || !nrow(cl) || !nrow(vl)) next
  for (wid in unique(cl$wave_id)) {
    daily <- merge(
      cl[wave_id == wid, .(mean_cr = mean(case_rate, na.rm = TRUE)), by = date],
      vl[wave_id == wid, .(total_visits = sum(total_visits, na.rm = TRUE)), by = date],
      by = "date", all = TRUE)
    if (nrow(daily) < 14) next
    setorder(daily, date)
    daily[, c_d := detrend(mean_cr)]
    daily[, v_d := detrend(total_visits)]
    n <- nrow(daily)
    for (lg in LAGS) {
      ok <- seq.int(max(1, 1 - lg), min(n, n - lg))
      if (length(ok) < 14) next
      x <- daily$v_d[ok]
      y <- daily$c_d[ok + lg]
      r <- suppressWarnings(cor(x, y, use = "pairwise.complete.obs"))
      ccf_rows[[length(ccf_rows) + 1]] <- data.table(
        state = state, wave_id = wid, lag = lg, cor = r,
        n_used = sum(!is.na(x) & !is.na(y))
      )
    }
  }
}

if (length(ccf_rows)) {
  ccf_dt <- rbindlist(ccf_rows)
  ccf_dt[, ci95 := 1.96 / sqrt(pmax(n_used, 1))]
  ccf_dt[, wave_label := mlabels[wave_id]]

  # Peak-lag table per (state, wave).
  peaks <- ccf_dt[, .SD[which.max(cor)], by = .(state, wave_id, wave_label)]
  peaks[, label := sprintf("peak ℓ = %d d   r = %.2f", lag, cor)]
  print(peaks[, .(state, wave_label, peak_lag_days = lag,
                    peak_cor = round(cor, 3), n_used)])

  ccf_csv <- file.path("outputs", "spatiotemporal_risk", "ccf_lag_scan.csv")
  dir.create(dirname(ccf_csv), recursive = TRUE, showWarnings = FALSE)
  fwrite(ccf_dt, ccf_csv)
  fwrite(peaks[, .(state, wave_id, wave_label, peak_lag_days = lag,
                     peak_cor = cor, n_used)],
         sub("\\.csv$", "_peaks.csv", ccf_csv))

  for (st in unique(ccf_dt$state)) {
    p <- ggplot(ccf_dt[state == st],
                aes(x = lag, y = cor, color = wave_label)) +
      geom_ribbon(aes(ymin = -ci95, ymax = ci95, fill = wave_label),
                  alpha = 0.10, color = NA) +
      geom_hline(yintercept = 0, color = "grey50") +
      geom_line(linewidth = 0.7) +
      geom_point(data = peaks[state == st],
                 aes(x = lag, y = cor), size = 2.5, shape = 21,
                 fill = "white", stroke = 1) +
      geom_text(data = peaks[state == st],
                aes(x = lag, y = cor, label = label),
                hjust = -0.1, vjust = -0.6, size = 3.2, show.legend = FALSE) +
      facet_wrap(~ wave_label, nrow = 1) +
      labs(title = paste(st, "— Cross-correlation: visits leading caserate"),
           subtitle = "Lag ℓ > 0 = visits precede cases by ℓ days (detrended series).",
           x = "Lag ℓ (days)", y = "Pearson cor",
           color = NULL, fill = NULL) +
      theme_bw() +
      theme(plot.title = element_text(face = "bold"),
            legend.position = "none",
            strip.background = element_rect(fill = "grey90"))
    print(p)
  }
}
```

