---
title: "06 - Origin-Destination Kernel rho(g,t)"
author: "Prince Oppong Boakye"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: show
    theme: flatly
---

```{r setup, include=FALSE}
source(here::here("helpers/_utils.R"))
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Overview

Computes contextual infection pressure rho(g,t) from OD mobility flows.
Each visit is weighted by incubation-weighted origin case rate (lags 1-7 days),
destination spillover rate, activity weight (visitors x dwell x distance decay),
and trip chain amplification. Outputs: aggregated rho per grid-day (for INLA)
and pair-level OD flows (for ABM coupling).

```{r init}
load_pipeline()
library(data.table)
library(sf)
library(arrow)

# Distance-kernel defaults (used as fallback if MLE fit fails). The fitted
# values per state-wave are written to the kernel diagnostics table.
OD_DECAY_M     <- 10000  # sigma (m), default
OD_DECAY_ALPHA <- 2      # power-law exponent, default (Meyer & Held 2014)

# Reporting-lag-aware incubation weights for COVID-19.
# Total infection -> case-record delay = 5d incubation + 1d testing + 1d reporting = 7d.
# Peak weight at t-7 reflects this; lighter mass at t-5..t-6 covers fast reporters
# and at t-1..t-2 covers known late-onset / repeat-exposure cases.
# Empirical CCF peaks from stage 04 (UT w2 lag-7, others 14-22d) support the
# lag-7 mass; longer-lag waves are absorbed by the RW2 temporal smooth in 09.
INCUBATION_WEIGHTS <- c(`7` = 0.30, `6` = 0.20, `5` = 0.20,
                        `4` = 0.10, `3` = 0.10, `2` = 0.05, `1` = 0.05)

# Quantile used for robust rho variant (Slater 2025 §5 future-work suggestion).
ROBUST_RHO_QUANTILE <- 0.5
```

# Helpers

```{r helpers}
load_cbg_modzcta_xw <- function(cfg) {
  rds <- file.path(cfg$cache_dir, "cbg_modzcta_crosswalk_NY.rds")
  if (file.exists(rds)) return(readRDS(rds))
  xw <- fread(cfg$nyc_crosswalk_path)
  xw[, cbg_geoid := as.character(cbg_geoid)]
  xw[, modzcta := as.character(modzcta)]
  dir.create(dirname(rds), showWarnings = FALSE, recursive = TRUE)
  saveRDS(xw, rds)
  xw
}

build_cbg_lookup <- function(cfg, state) {
  if (state == "NY") {
    xw <- load_cbg_modzcta_xw(cfg)
    data.table(cbg_geoid = standardize_cbg(xw$cbg_geoid),
               rate_geo_id = standardize_fips(xw$modzcta))
  } else {
    # Utah: county FIPS is first 5 chars of CBG GEOID
    cbg_sf <- st_read(get_cbg_path(state), quiet = TRUE)
    cbg_sf <- cbg_sf[substr(cbg_sf$GEOID, 1, 2) == cfg$ut_state_fips, ]
    data.table(cbg_geoid = standardize_cbg(cbg_sf$GEOID),
               rate_geo_id = substr(standardize_cbg(cbg_sf$GEOID), 1, 5))
  }
}

build_centroids <- function(cfg, state) {
  utm <- get_utm_crs(state)
  if (state == "NY") {
    cbg_sf <- st_read(cfg$ny_cbg_path, quiet = TRUE) |> st_transform(utm)
    cbg_sf <- cbg_sf[substr(cbg_sf$GEOID, 3, 5) %in% cfg$nyc_fips_3, ]
    xw <- load_cbg_modzcta_xw(cfg)
    cbg_sf$modzcta <- standardize_fips(xw$modzcta)[match(standardize_cbg(cbg_sf$GEOID), standardize_cbg(xw$cbg_geoid))]
    cbg_sf <- cbg_sf[!is.na(cbg_sf$modzcta), ]
    geo <- cbg_sf |> dplyr::group_by(modzcta) |> dplyr::summarise(geometry = sf::st_union(geometry))
    cc <- st_coordinates(st_centroid(geo))
    data.table(geo_id = standardize_fips(geo$modzcta), X = cc[,1], Y = cc[,2])
  } else {
    counties <- st_read(cfg$ut_counties_path, quiet = TRUE) |> st_transform(utm)
    fips_col <- grep("FIPS|fips|GEOID|geoid", names(counties), value = TRUE)[1]
    cc <- st_coordinates(st_centroid(counties))
    # UT counties shapefile stores FIPS as bare county code (1-57); case_rates
    # keys on 5-digit state+county (49xxx). Prepend state FIPS, then pad to 5.
    raw <- as.character(counties[[fips_col]])
    raw <- ifelse(nchar(raw) <= 3, paste0(cfg$ut_state_fips, sprintf("%03d", as.integer(raw))), raw)
    data.table(geo_id = standardize_fips(raw), X = cc[,1], Y = cc[,2])
  }
}

spillover_rates <- function(case_day, centroids, phi) {
  m <- merge(case_day, centroids, by = "geo_id", all = FALSE)
  if (nrow(m) < 2) { m[, contextual_rate := case_rate]; return(m[, .(geo_id, case_rate, contextual_rate)]) }
  dm <- as.matrix(dist(as.matrix(m[, .(X, Y)])))
  sm <- exp(-dm / phi); diag(sm) <- 0
  m[, contextual_rate := case_rate + as.vector(sm %*% m$case_rate)]
  m[, .(geo_id, case_rate, contextual_rate)]
}
```

# Main Processing

```{r process}
process_od_kernel <- function(wave_id, wave_months) {
  out_dir <- get_wave_dir(state, wave_id)
  rho_file <- file.path(out_dir, "tables", "od_kernel_rho.csv")
  dr <- get_wave_date_range(wave_months)
  n_days <- dr$n_days
  phi <- CONFIG$spillover_phi
  utm <- get_utm_crs(state)

  case_rates <- fread(file.path(out_dir, "tables", "case_rates.csv"))
  case_rates[, geo_id := standardize_fips(geo_id)]
  setkey(case_rates, geo_id, sim_day)

  cbg_lk    <- build_cbg_lookup(CONFIG, state)
  centroids <- build_centroids(CONFIG, state)

  grid_file <- file.path(out_dir, "intermediate", "grid_visits.feather")
  grid_dt <- as.data.table(arrow::read_feather(grid_file))
  grid_dt <- grid_dt[visits_total > 0]

  # map grids to geographic rate units
  if (!"rate_geo_id" %in% names(grid_dt) || sum(!is.na(grid_dt$rate_geo_id)) == 0) {
    gsf <- st_read(file.path(CONFIG$grids_dir, sprintf("grid_100m_%s.gpkg", state)), quiet = TRUE)
    gsf <- gsf[gsf$grid_id %in% grid_dt$grid_id, ]
    if (state == "UT" && "fips_5" %in% names(gsf)) {
      gm <- data.table(grid_id = gsf$grid_id, rate_geo_id = gsf$fips_5)
    } else {
      csf <- st_read(get_cbg_path(state), quiet = TRUE) |> st_transform(st_crs(gsf))
      csf <- csf[substr(csf$GEOID, 3, 5) %in% get_metro_fips(state, 3), ]
      gj <- st_join(st_centroid(gsf), csf[, "GEOID"], join = st_within, left = TRUE)
      gm <- data.table(grid_id = gj$grid_id, cbg_geoid = standardize_cbg(gj$GEOID))
      gm <- merge(gm, cbg_lk, by = "cbg_geoid", all.x = TRUE)
      gm <- gm[, .(grid_id, rate_geo_id)]
    }
    grid_dt <- merge(grid_dt, gm, by = "grid_id", all.x = TRUE)
  }
  grid_dt[, rate_geo_id := standardize_fips(rate_geo_id)]

  # load OD flows (wave-aggregated)
  od_file <- file.path(out_dir, "intermediate", "od_home_cbg.feather")
  od_dt <- as.data.table(arrow::read_feather(od_file))
  od_dt[, origin_cbg := standardize_cbg(home_cbg)]
  od_dt[, dest_cbg   := standardize_cbg(poi_cbg)]
  cbg_lk[, cbg_geoid := standardize_cbg(cbg_geoid)]

  od_dt <- merge(od_dt, cbg_lk, by.x = "origin_cbg", by.y = "cbg_geoid", all.x = TRUE)
  setnames(od_dt, "rate_geo_id", "origin_geo_id")
  od_dt <- merge(od_dt, cbg_lk, by.x = "dest_cbg", by.y = "cbg_geoid", all.x = TRUE)
  setnames(od_dt, "rate_geo_id", "dest_geo_id")

  # spatial join POIs to grid
  gsf <- st_read(file.path(CONFIG$grids_dir, sprintf("grid_100m_%s.gpkg", state)), quiet = TRUE) |> st_transform(utm)
  osf <- st_as_sf(od_dt[!is.na(poi_lon) & !is.na(poi_lat)], coords = c("poi_lon", "poi_lat"), crs = 4326) |> st_transform(utm)
  poi_xy <- st_coordinates(osf)
  osf$poi_utm_x <- poi_xy[,1]; osf$poi_utm_y <- poi_xy[,2]
  osf <- st_join(osf, gsf[, "grid_id"], left = FALSE)
  od <- as.data.table(st_drop_geometry(osf))
  od <- od[!is.na(origin_geo_id) & !is.na(dest_geo_id)]

  # activity weight
  od[, dwell_hours := pmin(median_dwell / 60, 8)]
  od[is.na(dwell_hours) | dwell_hours <= 0, dwell_hours := 0.5]
  od[, activity_weight := visitors * (1 + dwell_hours)]

  # distance decay -- fit power-law (sigma, alpha) from data; AIC vs Gaussian.
  cbg_cents <- st_read(get_cbg_path(state), quiet = TRUE) |> st_transform(utm)
  cbg_cents <- cbg_cents[substr(cbg_cents$GEOID, 3, 5) %in% get_metro_fips(state, 3), ]
  cc <- st_coordinates(st_centroid(cbg_cents))
  cbg_xy <- data.table(origin_cbg = standardize_cbg(cbg_cents$GEOID), home_X = cc[,1], home_Y = cc[,2])
  od <- merge(od, cbg_xy, by = "origin_cbg", all.x = TRUE)
  od[, od_distance_m := sqrt((poi_utm_x - home_X)^2 + (poi_utm_y - home_Y)^2)]

  k_fit <- fit_distance_kernel(od$od_distance_m, od$visitors)
  ## State-aware short/long threshold: visitor-weighted median trip distance,
  ## so rho_short / rho_long are comparable across NY (km-scale) and UT
  ## (10s-of-km scale). Falls back to 3 km when too few rows.
  short_long_thresh_m <- {
    ok <- is.finite(od$od_distance_m) & od$od_distance_m > 0 &
          is.finite(od$visitors) & od$visitors > 0
    if (sum(ok) >= 50) {
      d_ok <- od$od_distance_m[ok]; w_ok <- od$visitors[ok]
      ord <- order(d_ok); cw <- cumsum(w_ok[ord]) / sum(w_ok)
      d_ok[ord][which(cw >= 0.5)[1]]
    } else 3000
  }
  # use AIC-best form for distance weights
  if (k_fit$form == "tpowerlaw" && is.finite(k_fit$kappa)) {
    sigma_hat <- k_fit$sigma_tpl; alpha_hat <- k_fit$alpha_tpl; kappa_hat <- k_fit$kappa
  } else {
    sigma_hat <- k_fit$sigma;     alpha_hat <- k_fit$alpha;     kappa_hat <- NA_real_
  }
  message(sprintf("[%s/%s] OD kernel fit: form=%s sigma=%.0fm alpha=%.2f kappa=%s n=%d AIC_pl=%s AIC_tpl=%s AIC_g=%s",
                  state, wave_id, k_fit$form, sigma_hat, alpha_hat,
                  ifelse(is.na(kappa_hat),       "NA", sprintf("%.0fm", kappa_hat)),
                  k_fit$n,
                  ifelse(is.na(k_fit$aic_pl),    "NA", sprintf("%.0f", k_fit$aic_pl)),
                  ifelse(is.na(k_fit$aic_tpl),   "NA", sprintf("%.0f", k_fit$aic_tpl)),
                  ifelse(is.na(k_fit$aic_gauss), "NA", sprintf("%.0f", k_fit$aic_gauss))))
  diag_dir <- file.path(out_dir, "tables"); dir.create(diag_dir, showWarnings = FALSE, recursive = TRUE)
  fwrite(data.table(state = state, wave_id = wave_id,
                    sigma_m = sigma_hat, alpha = alpha_hat, kappa_m = kappa_hat,
                    sigma_gauss_m = k_fit$sigma_gauss,
                    aic_powerlaw = k_fit$aic_pl,
                    aic_tpowerlaw = k_fit$aic_tpl,
                    aic_gauss = k_fit$aic_gauss,
                    chosen_form = k_fit$form, n_pairs = k_fit$n,
                    short_long_thresh_m = short_long_thresh_m),
         file.path(diag_dir, "od_kernel_fit.csv"))

  od[!is.na(od_distance_m), distance_weight := (1 + od_distance_m / sigma_hat)^(-alpha_hat)]
  if (is.finite(kappa_hat)) {
    od[!is.na(od_distance_m), distance_weight := distance_weight * exp(-od_distance_m / kappa_hat)]
  }
  od[is.na(od_distance_m), distance_weight := 1]
  od[, activity_weight := activity_weight * distance_weight]
  od[, c("home_X", "home_Y") := NULL]

  # trip chain amplification (wave-level aggregation)
  chain_file <- file.path(out_dir, "intermediate", "trip_chains.feather")
  if (file.exists(chain_file) && "brand" %in% names(od)) {
    chains <- as.data.table(arrow::read_feather(chain_file))
    if (!"scope" %in% names(chains)) chains[, scope := "day"]
    bcs_day <- chains[scope == "day", .(chain_deg_day = sum(covisit_count)), by = source_brand]
    bcs_month <- chains[scope == "month", .(chain_deg_month = sum(covisit_count)), by = source_brand]
    bcs <- merge(bcs_day, bcs_month, by = "source_brand", all = TRUE)
    bcs[is.na(chain_deg_day), chain_deg_day := 0]
    bcs[is.na(chain_deg_month), chain_deg_month := 0]
    ## Normalise so the most-chained brand reaches boost = 1 exactly.
    ## (Old denominator `max(x + 1)` capped boost at 1 - 1/(max+1).)
    bcs[, boost_day   := chain_deg_day   / pmax(max(chain_deg_day),   1)]
    bcs[, boost_month := chain_deg_month / pmax(max(chain_deg_month), 1)]
    od <- merge(od, bcs[, .(source_brand, boost_day, boost_month)],
                by.x = "brand", by.y = "source_brand", all.x = TRUE)
    od[is.na(boost_day), boost_day := 0]
    od[is.na(boost_month), boost_month := 0]
    od[, activity_weight := activity_weight * (1 + 0.5 * boost_day + 0.25 * boost_month)]
    od[, c("boost_day", "boost_month") := NULL]
  }

  # crowding boost
  if ("avg_visit_density" %in% names(grid_dt)) {
    vd_grid <- grid_dt[, .(grid_id, avg_visit_density)]
    vd_med <- median(vd_grid$avg_visit_density[vd_grid$avg_visit_density > 0],
                     na.rm = TRUE)
    if (!is.na(vd_med) && vd_med > 0) {
      # Impute NA / non-positive cells to the median (delta_g = 1, neutral).
      # Earlier behaviour was to zero them, which silently muted the activity
      # weight wherever building-footprint coverage was poor.
      vd_grid[is.na(avg_visit_density) | avg_visit_density <= 0,
              avg_visit_density := vd_med]
      vd_grid[, crowding_factor := sqrt(pmin(avg_visit_density / vd_med, 5))]
      od <- merge(od, vd_grid[, .(grid_id, crowding_factor)],
                  by = "grid_id", all.x = TRUE)
      od[is.na(crowding_factor), crowding_factor := 1]
      od[, activity_weight := activity_weight * crowding_factor]
      od[, crowding_factor := NULL]
    }
  }

  # ABM pair-level flows
  pcols <- intersect(c("origin_cbg", "origin_geo_id", "grid_id", "dest_geo_id", "poi_lon", "poi_lat",
                        "visitors", "median_dwell", "dwell_hours", "activity_weight", "od_distance_m",
                        "brand", "naics_code"), names(od))
  pairs <- od[, ..pcols]
  setnames(pairs, "grid_id", "dest_grid_id")
  pairs_agg <- pairs[, .(visitors = sum(visitors),
                          dwell_hours = weighted.mean(dwell_hours, visitors, na.rm = TRUE),
                          activity_weight = sum(activity_weight),
                          od_distance_m = weighted.mean(od_distance_m, visitors, na.rm = TRUE),
                          n_pois = .N),
                      by = .(origin_cbg, origin_geo_id, dest_grid_id, dest_geo_id)]
  pairs_agg[, travel_prob := activity_weight / sum(activity_weight), by = origin_cbg]
  pairs_agg[, transmission_weight := visitors * dwell_hours]
  od_pairs_export <- copy(pairs_agg)
  rm(pairs, pairs_agg); gc(verbose = FALSE)

  # daily rho computation (incubation-weighted origin rates) over 90 continuous days
  rho_list <- vector("list", n_days)
  for (day in seq_len(n_days)) {
    lag_days <- as.integer(names(INCUBATION_WEIGHTS))
    lag_rates <- rbindlist(lapply(seq_along(lag_days), function(k) {
      sd <- day - lag_days[k]
      if (sd < 1) sd <- 1L  # clamp at wave start (no prior data)
      cr <- case_rates[sim_day == sd, .(case_rate = mean(case_rate)), by = .(origin_geo_id = geo_id)]
      cr[, w := INCUBATION_WEIGHTS[k]]
      cr
    }))
    oday <- lag_rates[, .(origin_rate = sum(case_rate * w) / sum(w)), by = origin_geo_id]
    dday <- case_rates[sim_day == max(1, day - 1), .(case_rate = mean(case_rate)), by = geo_id]
    dctx <- spillover_rates(dday, centroids, phi)
    setnames(dctx, "geo_id", "dest_geo_id")

    dod <- copy(od)
    dod <- merge(dod, oday, by = "origin_geo_id", all.x = TRUE)
    dod <- merge(dod, dctx[, .(dest_geo_id, contextual_rate)], by = "dest_geo_id", all.x = TRUE)
    dod[is.na(origin_rate), origin_rate := 0]
    dod[is.na(contextual_rate), contextual_rate := 0]
    dod[, visit_risk := origin_rate * contextual_rate]
    ## State-aware short/long split: median trip distance (computed once below)
    ## so `rho_short` / `rho_long` are comparable across NY (km-scale) and UT
    ## (10s-of-km scale).
    dod[, is_short := !is.na(od_distance_m) & od_distance_m < short_long_thresh_m]

    # Robust rho: quantile-weighted (median) rather than activity-weighted mean.
    # Insensitive to early-pandemic case-count spikes (Slater 2025 §5 future work).
    dod[, contrib := visit_risk * activity_weight]
    rho_list[[day]] <- dod[, .(
      rho_raw   = sum(contrib, na.rm = TRUE),
      rho_robust = if (.N > 0)
                     stats::quantile(contrib[contrib > 0], ROBUST_RHO_QUANTILE,
                                     na.rm = TRUE, names = FALSE) * .N
                   else 0,
      rho_short = sum(visit_risk[is_short] * activity_weight[is_short], na.rm = TRUE),
      rho_long  = sum(visit_risk[!is_short] * activity_weight[!is_short], na.rm = TRUE),
      exposure_import  = sum(activity_weight * origin_rate, na.rm = TRUE),
      exposure_gravity = sum(activity_weight * origin_rate / pmax(od_distance_m / 1000, 0.1)^1.5, na.rm = TRUE),
      total_activity = sum(activity_weight), n_od_pairs = .N, total_visitors = sum(visitors)
    ), by = grid_id][, sim_day := day]
  }
  rho_dt <- rbindlist(rho_list, fill = TRUE)
  all_gd <- CJ(grid_id = unique(grid_dt$grid_id), sim_day = 1:n_days)
  rho_dt <- merge(all_gd, rho_dt, by = c("grid_id", "sim_day"), all.x = TRUE)
  for (cl in c("rho_raw", "rho_robust", "rho_short", "rho_long",
               "exposure_import", "exposure_gravity"))
    rho_dt[is.na(get(cl)), (cl) := 0]
  rho_dt[, rho := rho_raw]

  # normalization (single wave-wide pass — no inter-month calibration)
  gmean <- mean(rho_dt$rho[rho_dt$rho > 0], na.rm = TRUE)
  if (is.na(gmean) || gmean <= 0) gmean <- 1
  rho_dt[, rho_norm := pmax(rho / gmean, 0.01)]
  pmean <- mean(rho_dt$rho_norm)
  if (abs(pmean - 1) > 0.05) rho_dt[, rho_norm := rho_norm / pmean]
  rho_dt[, log_rho := log(rho_norm)]

  for (sc in c("rho_short", "rho_long", "rho_robust")) {
    sm <- mean(rho_dt[[sc]][rho_dt[[sc]] > 0], na.rm = TRUE)
    if (is.na(sm) || sm <= 0) sm <- 1
    rho_dt[, (paste0(sc, "_norm")) := pmax(get(sc) / sm, 0.01)]
  }
  rho_dt[, log_rho_short := log(rho_short_norm)]
  rho_dt[, log_rho_long := log(rho_long_norm)]
  rho_dt[, log_rho_robust := log(rho_robust_norm)]
  rho_dt[, log_exposure_import := log1p(exposure_import)]
  rho_dt[, log_exposure_gravity := log1p(fifelse(is.na(exposure_gravity), 0, exposure_gravity))]

  rho_dt[, kernel_mode := "TRUE_OD"]
  fwrite(rho_dt, rho_file)

  # ABM pairs (wave-level).  Schema:
  #   origin_cbg, origin_geo_id, dest_grid_id, dest_geo_id,
  #   visitors, dwell_hours, activity_weight, od_distance_m, n_pois,
  #   travel_prob (per-origin), transmission_weight (= visitors * dwell_hours)
  # Not consumed by the INLA pipeline; preserved for downstream ABM coupling
  # (see thesis Sec. 1 -- "OD pair-level output ... preserves the
  # infrastructure for future ABM coupling").
  pairs_file <- file.path(out_dir, "network", "od_pairs.feather")
  dir.create(dirname(pairs_file), showWarnings = FALSE, recursive = TRUE)
  arrow::write_feather(od_pairs_export, pairs_file)

  # spatial lag for 07
  rho_grid <- rho_dt[, .(rho_mean = mean(rho_norm), rho_sd = sd(rho_norm),
                          log_rho_mean = mean(log_rho)), by = grid_id]
  arrow::write_feather(rho_grid, file.path(out_dir, "intermediate", "spatial_lag.feather"))

  # persist geo mapping
  gv <- as.data.table(arrow::read_feather(grid_file))
  if (!"rate_geo_id" %in% names(gv)) {
    gv <- merge(gv, grid_dt[, .(grid_id, rate_geo_id)], by = "grid_id", all.x = TRUE)
    arrow::write_feather(gv, grid_file)
  }

  list(wave_id = wave_id, n_griddays = nrow(rho_dt),
       rho_range = round(range(rho_dt$rho_norm), 3),
       rho_nonzero_pct = round(100 * sum(rho_dt$rho > 0) / nrow(rho_dt), 1),
       od_pairs_n = nrow(od_pairs_export))
}
```

# Run All

```{r run}
results_dt <- rbindlist(run_waves_by_state(process_od_kernel), fill = TRUE)
fwrite(results_dt, file.path(CONFIG$output_dir, "od_kernel_run_summary.csv"))
print(results_dt)
```

## Month-Seam Diagnostic

```{r month_seams, fig.width=10, fig.height=6}
## Empirical check that month-boundary discontinuities (Advan monthly
## aggregation -> uniform within-month allocation) do not introduce visible
## step changes in log_rho. Vertical lines mark the 1st of each month.
library(ggplot2)
seam_dt <- rbindlist(lapply(seq_len(nrow(WAVE_RUNS)), function(i) {
  st  <- WAVE_RUNS$state[i]; wid <- WAVE_RUNS$wave_id[i]
  rho_path <- file.path(get_wave_dir(st, wid), "tables", "od_kernel_rho.csv")
  if (!file.exists(rho_path)) return(NULL)
  cr_path  <- file.path(get_wave_dir(st, wid), "tables", "case_rates.csv")
  dates    <- fread(cr_path)[, .(sim_day, date = as.Date(date))][, .SD[1], by = sim_day]
  dr       <- fread(rho_path)[, .(log_rho_mean = mean(log_rho, na.rm = TRUE)), by = sim_day]
  merge(dr, dates, by = "sim_day")[, `:=`(state = st, wave_id = wid)]
}), fill = TRUE)
if (nrow(seam_dt)) {
  seams <- seam_dt[, .(month_start = unique(as.Date(format(date, "%Y-%m-01")))[-1L]),
                   by = .(state, wave_id)]
  fwrite(seam_dt, file.path(CONFIG$output_dir, "log_rho_daily_means.csv"))
  print(ggplot(seam_dt, aes(date, log_rho_mean)) +
    geom_line(linewidth = 0.4) +
    geom_vline(data = seams, aes(xintercept = month_start),
               linetype = "dashed", colour = "red", alpha = 0.5) +
    facet_wrap(state ~ wave_id, scales = "free", ncol = 3) +
    labs(title = "Daily mean log(rho) with month-boundary markers",
         subtitle = "Red dashes = 1st of each calendar month within wave window",
         x = NULL, y = "mean log(rho) across grids") +
    theme_ts())
}
```

```{r session}
sessionInfo()
```
