---
title: "08b — Change-of-Support Disaggregation Diagnostics"
subtitle: "Visualizing the Poisson disaggregation from MODZCTA to 100 m grid (NYC)"
output: html_document
---

This notebook visualises the change-of-support step from `08a_build_model_data.Rmd`.
It checks the population offset, grid-level disaggregation, Poisson calibration,
zero inflation, MODZCTA round-trip aggregation, daily totals, and within-zone
allocation variation.

```{r setup, include=FALSE}
source(here::here("helpers/_utils.R"))
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      fig.width = 12, fig.height = 8)
```

```{r setup, include=FALSE}
source(here::here("helpers/_utils.R"))
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      fig.width = 12, fig.height = 8)
```

```{r libraries}
load_pipeline()
library(data.table)
library(sf)
library(ggplot2)
library(patchwork)
library(viridis)
library(ggspatial)
library(scales)
library(knitr)

states <- unique(WAVE_RUNS$state)
```

```{r load_data}
all_state_data <- list()
for (state in states) {
  .this_state <- state
  st_waves <- WAVE_RUNS[which(WAVE_RUNS$state == .this_state), ]
  wave_ids <- st_waves$wave_id
  grid_sf0 <- st_read(file.path(CONFIG$grids_dir, sprintf("grid_100m_%s.gpkg", state)), quiet = TRUE)
  all_md <- setNames(lapply(wave_ids, function(w) {
    fp <- file.path(get_wave_dir(state, w), "models", "model_data_gt.rds")
    readRDS(fp)
  }), wave_ids)
  modzcta_sf <- NULL
  if (state == "NY") {
    mz_files <- list.files(file.path("data", "MODZCTA"),
                           pattern = "\\.shp$", full.names = TRUE)
    if (length(mz_files) >= 1) {
      modzcta_sf <- st_read(mz_files[1], quiet = TRUE) |>
        st_transform(st_crs(grid_sf0))
      # Normalise the zone id column to 'modzcta' (case- and name-tolerant)
      mz_col <- intersect(c("modzcta", "MODZCTA", "MODZCTA_CODE", "zcta"),
                          names(modzcta_sf))[1]
      if (!is.na(mz_col) && mz_col != "modzcta")
        names(modzcta_sf)[names(modzcta_sf) == mz_col] <- "modzcta"
      if ("modzcta" %in% names(modzcta_sf))
        modzcta_sf$modzcta <- as.character(modzcta_sf$modzcta)
    }
  }
  # Keep the months field name for downstream chunks.
  all_state_data[[state]] <- list(grid_sf0 = grid_sf0, all_md = all_md,
                                  months = wave_ids, modzcta_sf = modzcta_sf)
}
poly_color <- NA
poly_lw    <- 0
```

# 1. Expected Population E

The grid-level expected population $E$ combines census population with building
density weights so that each grid cell's share of the MODZCTA population is
proportional to its built-environment footprint.

```{r E_map, fig.height=7}
for (state in states) {
  sd <- all_state_data[[state]]
  grid_sf0 <- sd$grid_sf0; all_md <- sd$all_md
  gt0 <- as.data.table(all_md[[1]]$gt_dt)
  grid_E <- gt0[, .(E = E[1]), by = grid_id]
  grid_plot_E <- merge(grid_sf0, grid_E, by = "grid_id")

  print(ggplot(grid_plot_E) +
    geom_sf(aes(fill = E), color = NA, lwd = 0) +
    scale_fill_viridis_c(option = "mako", name = "E", trans = "log1p",
                         breaks = c(0, 10, 50, 200, 1000), labels = comma) +
    map_extras() +
    labs(title = paste(state, "\u2014 Expected Population E per Grid Cell"),
         subtitle = "log(1+E) scale \u2014 WorldPop 100 m population (02a)") +
    theme_map())
}
```

# 2. Cumulative Y per Wave

For each grid cell and day, the expected Poisson mean is

$$\lambda_{g,t} = \max\!\Bigl(\frac{r_{a,t} \times E_g}{10^5},\; 10^{-6}\Bigr)$$

Here $r_{a,t}$ is the reporting-area daily incidence rate per 100,000 from the
external case-rate table. The area-day integer total

$$N_{a,t} = \text{round}\bigl(\sum_{g \in a} \lambda_{g,t}\bigr)$$

is disaggregated via

$$\mathbf{Y}_{a,t} \sim \text{Multinomial}(N_{a,t},\; \mathbf{p}_a),\qquad
p_g = \frac{E_g}{\sum_{g \in a} E_g}.$$

With $w_g = 1$, allocation depends only on the grid population offset.
Re-aggregating the grid-level counts recovers the observed area-day total
exactly, so no cases are created or lost.

```{r Y_maps, fig.height=12}
for (state in states) {
  sd <- all_state_data[[state]]
  grid_sf0 <- sd$grid_sf0; all_md <- sd$all_md; months <- sd$months
  plots <- lapply(months, function(m) {
    gt <- as.data.table(all_md[[m]]$gt_dt)
    grid_agg <- gt[, .(Y_total = sum(Y, na.rm = TRUE)), by = grid_id]
    gp <- merge(grid_sf0, grid_agg, by = "grid_id")
    ggplot(gp) +
      geom_sf(aes(fill = Y_total), color = NA, lwd = 0) +
      scale_fill_viridis_c(option = "rocket", name = expression(Sigma * Y),
                           trans = "log1p") +
      labs(title = paste0(state, " ", m)) +
      map_extras() + theme_map()
  })
  print(wrap_plots(plots, ncol = min(3, length(months))) +
    plot_annotation(title = paste(state, "\u2014 Cumulative Disaggregated Counts (Y)"),
                    subtitle = "Sum over all days in the wave; log(1+x) scale"))
}
```

# 3. Lambda vs Y Scatter

A well-calibrated disaggregation should show Y values scattered around the
1:1 line with variance roughly equal to the mean (Poisson property).

```{r lambda_y_scatter, fig.height=6}
for (state in states) {
  sd <- all_state_data[[state]]
  all_md <- sd$all_md; months <- sd$months
  samp_month <- months[1]
  gt_samp <- as.data.table(all_md[[samp_month]]$gt_dt)
  if (!"lambda_cases" %in% names(gt_samp)) {
    next
  }
  set.seed(1)
  idx <- if (nrow(gt_samp) > 50000) sample(nrow(gt_samp), 50000) else seq_len(nrow(gt_samp))
  samp <- gt_samp[idx, .(lambda_cases, Y)]
  print(ggplot(samp, aes(x = lambda_cases, y = Y)) +
    geom_point(alpha = 0.05, size = 0.3, color = "steelblue") +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "red") +
    scale_x_log10(labels = comma) + scale_y_log10(labels = comma) +
    labs(title = paste0(state, " — λ vs Y (", samp_month, ")"),
         subtitle = "Dashed red = 1:1 line",
         x = expression(lambda[cases]), y = "Y (Poisson draw)") +
    theme_minimal(base_size = 13))
}
```

# 4. Distribution of Y and Lambda

```{r histograms, fig.height=5}
for (state in states) {
  sd <- all_state_data[[state]]
  all_md <- sd$all_md; months <- sd$months
  gt_all <- rbindlist(lapply(months, function(m) {
    dt <- as.data.table(all_md[[m]]$gt_dt)
    cols <- intersect(c("lambda_cases", "Y"), names(dt))
    dt <- dt[, ..cols]
    dt[, month := m]
  }))
  p1 <- ggplot(gt_all[Y > 0], aes(x = Y)) +
    geom_histogram(bins = 60, fill = "steelblue", color = "white", linewidth = 0.2) +
    scale_x_log10(labels = comma) +
    facet_wrap(~month, scales = "free_y") +
    labs(title = paste(state, "— Distribution of Y (non-zero)"), x = "Y", y = "Count") +
    theme_minimal(base_size = 11)
  if ("lambda_cases" %in% names(gt_all)) {
    p2 <- ggplot(gt_all[lambda_cases > 1e-5], aes(x = lambda_cases)) +
      geom_histogram(bins = 60, fill = "darkorange", color = "white", linewidth = 0.2) +
      scale_x_log10(labels = comma) +
      facet_wrap(~month, scales = "free_y") +
      labs(title = paste(state, "— Distribution of λ (non-trivial)"), x = expression(lambda), y = "Count") +
      theme_minimal(base_size = 11)
    print(p1 / p2)
  } else {
    print(p1)
  }
}
```

# 5. Zero Inflation

Many grid cells will have $Y = 0$ because their Poisson rate is very small.
This is expected: most fine-grained cells have negligible population exposure.

```{r zero_inflation, fig.height=4}
for (state in states) {
  sd <- all_state_data[[state]]
  all_md <- sd$all_md; months <- sd$months
  zi <- rbindlist(lapply(months, function(m) {
    gt <- as.data.table(all_md[[m]]$gt_dt)
    has_lambda <- "lambda_cases" %in% names(gt)
    data.table(month = m, n_obs = nrow(gt),
               pct_zero_Y = 100 * mean(gt$Y == 0),
               pct_zero_lambda = if (has_lambda) 100 * mean(gt$lambda_cases <= 1e-6) else NA_real_,
               mean_lambda = if (has_lambda) mean(gt$lambda_cases) else NA_real_,
               mean_Y = mean(gt$Y))
  }))
  print(kable(zi, digits = 2, caption = paste(state, "— Zero-inflation summary")))
  p <- ggplot(zi, aes(x = month)) +
    geom_col(aes(y = pct_zero_Y), fill = "steelblue", width = 0.6)
  if (!all(is.na(zi$pct_zero_lambda)))
    p <- p + geom_point(aes(y = pct_zero_lambda), color = "red", size = 3)
  p <- p + labs(title = paste(state, "— Percentage of Zero Observations"),
                subtitle = "Bars = Y == 0 | Red dots = λ ≤ 1e-6 (grid states only)",
                x = "Month", y = "% Zero") +
    theme_minimal(base_size = 13)
  print(p)
}
```

# 6. MODZCTA Round-Trip: Observed vs Disaggregated

The most stringent diagnostic of the change-of-support step is whether
re-aggregating the grid-level $Y_{g,t}$ back to MODZCTA recovers the
**externally observed** zone-day case totals from `case_rates.csv` (02b).

For each MODZCTA $a$ and day $t$ we compute

$$
\widehat C_{a,t} \;=\; \sum_{g\in a} Y_{g,t}\quad\text{vs}\quad
C_{a,t} \;=\; \frac{r^{\,\text{spill}}_{a,t}}{10^{5}}\cdot\frac{1}{7}\cdot\sum_{g\in a} E_{g}
$$

where $r^{\,\text{spill}}_{a,t}$ is the borough-spillover-adjusted weekly
incidence per 100k from 02b. A well-behaved disaggregation gives a tight
1:1 cloud and a residual map without spatial bias.

```{r modzcta_roundtrip, fig.height=7}
for (state in states) {
  sd <- all_state_data[[state]]
  all_md <- sd$all_md; months <- sd$months
  if (state != "NY") next  # only NY has MODZCTA disaggregation diagnostics

  # ---- per-wave round-trip ----
  rt_per_wave <- lapply(months, function(m) {
    gt <- as.data.table(all_md[[m]]$gt_dt)
    case_dt <- fread(file.path(get_wave_dir(state, m), "tables", "case_rates.csv"))
    case_dt[, geo_id := as.character(geo_id)]
    # Observed expected cases per MODZCTA-day from external 02b table.
    # Deduplicate E by grid before summing across days.
    grid_E  <- unique(gt[, .(grid_id, geo_id, E)])
    zone_E  <- grid_E[, .(E_total = sum(E)), by = geo_id]
    obs <- merge(case_dt[, .(geo_id, sim_day, case_rate_spillover)],
                 zone_E, by = "geo_id")
    obs[, C_obs := case_rate_spillover / 1e5 * E_total / 7]
    # Round-trip from grid
    rt <- gt[, .(C_hat = sum(Y)), by = .(geo_id, sim_day)]
    rt <- merge(rt, obs[, .(geo_id, sim_day, C_obs)],
                by = c("geo_id", "sim_day"))
    rt[, wave := m]
    rt
  })
  rt_all <- rbindlist(rt_per_wave)

  # ---- (a) per-(zone,day) scatter ----
  print(ggplot(rt_all, aes(x = C_obs, y = C_hat)) +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "red") +
    geom_point(alpha = 0.05, size = 0.4, color = "steelblue") +
    geom_smooth(method = "gam", formula = y ~ s(x, bs = "cs"), se = FALSE,
                color = "black", linewidth = 0.6) +
    facet_wrap(~ wave, ncol = min(3, length(months))) +
    scale_x_log10(labels = comma) + scale_y_log10(labels = comma) +
    labs(title = paste(state,
         "\u2014 MODZCTA\u00d7day round-trip: disaggregated \u03a3Y vs observed expected cases"),
         subtitle = "Each point = one (MODZCTA, day); red = 1:1; black = GAM smoother",
         x = expression(C[a*","*t] ~ "(observed, from 02b)"),
         y = expression(hat(C)[a*","*t] ~ "(\u03a3 grid Y)")) +
    theme_minimal(base_size = 12))

  # ---- (b) per-MODZCTA totals (collapse over days) ----
  rt_zone <- rt_all[, .(C_obs = sum(C_obs), C_hat = sum(C_hat)),
                    by = .(geo_id, wave)]
  rt_zone[, rel_err := (C_hat - C_obs) / pmax(C_obs, 1)]

  # Goodness-of-fit metrics per wave
  metrics <- rt_zone[, .(
      n_zones = .N,
      median_rel_err = scales::percent(median(rel_err, na.rm = TRUE), 0.1),
      iqr_rel_err    = paste0(
        scales::percent(quantile(rel_err, 0.25, na.rm = TRUE), 0.1), " / ",
        scales::percent(quantile(rel_err, 0.75, na.rm = TRUE), 0.1)),
      pearson_r = round(cor(C_hat, C_obs, use = "complete.obs"), 4),
      bias_pct  = scales::percent(sum(C_hat) / sum(C_obs) - 1, 0.01)
    ), keyby = wave]
  print(kable(metrics,
    caption = paste(state,
      "\u2014 round-trip goodness-of-fit per wave (over MODZCTA totals)")))

  # ---- (c) residual choropleth ----
  if (!is.null(sd$modzcta_sf)) {
    res_plots <- lapply(months, function(m) {
      d <- rt_zone[wave == m]
      mz <- merge(sd$modzcta_sf, d, by.x = "modzcta", by.y = "geo_id")
      lim <- min(0.5, max(abs(mz$rel_err), na.rm = TRUE))
      ggplot(mz) +
        geom_sf(aes(fill = pmin(pmax(rel_err, -lim), lim)),
                color = "grey70", lwd = 0.05) +
        scale_fill_gradient2(low = "#2c7bb6", mid = "white", high = "#d7191c",
                             midpoint = 0, limits = c(-lim, lim),
                             oob = scales::squish,
                             name = "Relative\nresidual",
                             labels = scales::percent_format(1)) +
        labs(title = m) + theme_map()
    })
    print(wrap_plots(res_plots, ncol = min(3, length(months))) +
      plot_annotation(
        title = paste(state, "\u2014 Per-MODZCTA round-trip residual (\u03a3Y \u2212 C_obs)/C_obs"),
        subtitle = "Blue = under-allocated, red = over-allocated; clipped at \u00b150%"))
  }
}
```

# 7. Temporal Daily Totals

Day-by-day comparison of the total disaggregated count versus the total
Poisson rate across all grid cells.

```{r daily_totals, fig.height=6}
for (state in states) {
  sd <- all_state_data[[state]]
  all_md <- sd$all_md; months <- sd$months
  daily_plots <- lapply(months, function(m) {
    gt <- as.data.table(all_md[[m]]$gt_dt)
    has_lambda <- "lambda_cases" %in% names(gt)
    daily <- if (has_lambda) {
      gt[, .(Y_total = sum(Y), lambda_total = sum(lambda_cases)), by = sim_day]
    } else {
      gt[, .(Y_total = sum(Y)), by = sim_day]
    }
    setorder(daily, sim_day); daily[, day_idx := .I]
    p <- ggplot(daily, aes(x = day_idx))
    if (has_lambda)
      p <- p + geom_line(aes(y = lambda_total, color = "Expected (λ)"), linewidth = 0.8)
    p <- p +
      geom_line(aes(y = Y_total, color = "Observed (Y)"), linewidth = 0.5, alpha = 0.7) +
      scale_color_manual(values = c("Expected (λ)" = "darkorange",
                                    "Observed (Y)" = "steelblue")) +
      scale_y_continuous(labels = comma) +
      labs(title = m, x = "Day", y = "Total Count", color = NULL) +
      theme_minimal(base_size = 11) + theme(legend.position = "bottom")
    p
  })
  print(wrap_plots(daily_plots, ncol = 1) +
    plot_annotation(title = paste(state, "— Daily Totals: Y vs λ"),
                    subtitle = "Y tracks λ closely — Poisson noise is small relative to total"))
}
```

# 8. Spatial Allocation Effectiveness

To verify that building density meaningfully differentiates allocation,
we compare the coefficient of variation (CV) of $\lambda$ across grid cells
within each MODZCTA: higher CV indicates more spatial differentiation.

```{r allocation_cv, fig.height=6}
for (state in states) {
  sd <- all_state_data[[state]]
  all_md <- sd$all_md; months <- sd$months
  gt0 <- as.data.table(all_md[[1]]$gt_dt)
  if (!"lambda_cases" %in% names(gt0)) {
    next
  }
  cv_dt <- gt0[lambda_cases > 1e-6,
               .(cv_lambda = sd(lambda_cases) / mean(lambda_cases), n_grids = .N),
               by = geo_id]
  cv_dt <- cv_dt[n_grids >= 5]
  geo_label <- if (state == "NY") "MODZCTA" else "County"
  print(ggplot(cv_dt, aes(x = cv_lambda)) +
    geom_histogram(bins = 40, fill = "steelblue", color = "white") +
    labs(title = paste(state, "\u2014 Within-zone CV of \u03bb"),
         subtitle = paste0("Higher CV = more spatial differentiation (", months[1], ")"),
         x = paste0("CV(\u03bb) within ", geo_label), y = paste("Number of", geo_label, "s")) +
    theme_minimal(base_size = 13))
}
```
