---
title: "09 - Unified Component Model (clean, shareable)"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: show
    theme: flatly
---


## Overview

This cleaned, shareable version of the `09_fit_model` pipeline is a minimal, publication-ready template. It removes personal debugging narratives, cluster-specific anecdotes, and large in-line run logs while preserving the core pipeline structure and the points where users should provide data and configuration.

Required inputs

- `CONFIG` object (paths to `grids_dir`, `output_dir`, and other resources).
- `helpers/_utils.R` and `helpers/_adjacency.R` should be available (this template assumes core helper functions such as `fit_model()`, `extract_model_outputs()`, `build_formulas()` are defined there).

Minimal run options (via environment variables)

- `fit_state`, `fit_wave` — restrict to a single state/wave
- `fit_num_threads` — INLA threading string (e.g. "1:2" or "4:9")
- `fit_force_refit` — set to `1` to force all variants to re-fit
- `fit_force_warmstart` — set to `1` to allow warm-starting from related models
- `fit_full_cpo`, `fit_full_quality` — toggles for evaluation/integration strategies
- `fit_n_st_clusters` — override `N_ST_CLUSTERS` when desired

Note: This file is a compact, shareable template. For the full, production pipeline with all optimizations and data-specific handling, see `pipeline/09_fit_model.Rmd`.

```{r setup, include=FALSE}
source(here::here("helpers/_utils.R"))
source(here::here("helpers/_adjacency.R"))
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

# Simple timestamped logger used in the template
log_msg <- function(...) {
  cat(sprintf("[%s]", format(Sys.time(), "%H:%M:%S")), " ", ..., "\n", sep = "")
}
```

## Configuration

```{r config}
# Minimal, sensible defaults for shareable template
N_ST_CLUSTERS <- as.integer(Sys.getenv("fit_n_st_clusters", 40))
.fit_num_threads <- Sys.getenv("fit_num_threads", "1:2")
.fit_full_cpo <- identical(Sys.getenv("fit_full_cpo"), "1")
.fit_full_quality <- identical(Sys.getenv("fit_full_quality"), "1")
# Other CONFIG entries expected to be provided externally (project-specific)
```

## Example: Cleaned Save & Comparison Snippet

This chunk demonstrates a concise, shareable approach to persisting fit outputs and performing model comparison. It intentionally omits verbose run-time narrative and cluster-specific advice.

```{r fit-snippet}
# -- Within the per-model loop (simplified) --
# Assume: 'fit' is the returned object from fit_model();
#         'comp' is a data.table accumulating comparison rows;
#         'ckpt_out', 'ckpt_comp', 'ckpt_conv', 'ckpt_dir_t' are defined paths.

if (fit$status == "OK") {
  # Extract and persist model outputs
  mo <- extract_model_outputs(fit$result, df, subset$grid_index, mn, E_vec)
  saveRDS(mo, ckpt_out)

  # Save posterior hyperparameters (for optional warm-starting)
  ws_theta <- fit$result$mode$theta
  if (is.null(ws_theta) && !is.null(fit$result$summary.hyperpar)) {
    ws_theta <- fit$result$summary.hyperpar$mode
  }
  if (!is.null(ws_theta) && length(ws_theta) > 0) {
    saveRDS(ws_theta, file.path(ckpt_dir_t, paste0(mn, "_mode.rds")))
    log_msg("Saved theta mode for", mn)
  }

  # Save latent posterior mean (optional warm-start vector)
  ws_x <- fit$result$mode$x
  if (!is.null(ws_x) && length(ws_x) > 0) {
    saveRDS(ws_x, file.path(ckpt_dir_t, paste0(mn, "_xmode.rds")))
    log_msg("Saved x-mode for", mn)
  }

  rm(mo)
}

# Basic memory cleanup
rm(fit)
gc(full = TRUE)
```

```{r comparison-snippet}
# Reload extracted outputs for OK models (example)
ok_models <- comp[Status == "OK", Model]
outputs <- list()
for (mn in ok_models) {
  pth <- file.path(ckpt_dir_m, paste0(mn, ".rds"))
  if (file.exists(pth)) outputs[[mn]] <- readRDS(pth)
}

# Model comparison: compute deltas defensively
ok <- comp[Status == "OK"]
if (nrow(ok) > 0) {
  ok[, dWAIC := if (all(is.na(WAIC))) NA_real_ else WAIC - min(WAIC, na.rm = TRUE)]
  ok[, dDIC := if (all(is.na(DIC))) NA_real_ else DIC - min(DIC, na.rm = TRUE)]
  comp[Status == "OK", `:=`(dWAIC = ok$dWAIC, dDIC = ok$dDIC)]
  print(ok[order(WAIC)])
  if (!all(is.na(ok$WAIC))) {
    best_model <- ok[which.min(WAIC), Model]
  } else {
    best_model <- NA_character_
  }
}
```

## Minimal save of final artifacts

```{r save-final}
# Create a compact final object for downstream steps
final <- list(
  state = state, wave_id = wave_id, best_model = best_model,
  comparison = comp, n_obs = nrow(df)
)
saveRDS(final, file.path(out_dir, "models", "model_results_unified.rds"))
fwrite(comp, file.path(out_dir, "tables", "model_comparison_unified.csv"))
```

## Notes for users

```{r}
# End of template
```
